---
title: "Create Scores Overview"
author: "Annika Kuelpmann"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = TRUE)

```

# Preparations 
## Load Packages
```{r packages}
#| message: false
#| warning: false

# load R packages 
library(knitr)
library(tidyverse)
library(readxl)
library(writexl)
library(conflicted)
library(kableExtra)

# declare preferences regarding conflicts 
conflicts_prefer(dplyr::filter)

```

## Load Data 
```{r housekeeping}

# read scoring master file
data_scoring <- read_xlsx("../../data/raw/metamethod review/master coding/codebook_scoring_master.xlsx") 

# read monotonic scores file
# this is a file where we identified scores that can be transformed into each other while preserving ranks
monotonic <- read_xlsx("../../data/raw/metamethod review/master coding/monotonic_scores.xlsx")

```

# Distinct Scores
This code creates an object with the distinct scores as described in the IGT articles.

```{r overview raw}

# create object with distinct scores
scores_overview <- data_scoring %>%
  distinct(score_id, .keep_all=TRUE)

count_scores <- nrow(scores_overview)

count_scores |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Monotonic Scores
Some scores in the sample are described and reported as if they were distinct from each other, but they can be transformed into each other while perfectly preserving ranks. Here, we collapse these scores to prevent an overestimation of the total number of distinct scores. 

```{r overview monotonic}
# join info on monotonic scores to overview of all scores
scores_monotonic <- scores_overview %>% 
  left_join(monotonic, by = "score_id") %>% 
  mutate(monotonic_score_id = ifelse(is.na(new_score_id), score_id, new_score_id)) %>% 
  distinct(monotonic_score_id, .keep_all=TRUE)

# align score descriptions and formulas of monotonic scores 
scores_monotonic <- scores_monotonic %>%
 mutate(
    score_description = case_when(
        str_detect(score_description, fixed("[(C+D)-(A+B)]")) |
        str_detect(score_description, fixed("(C+D)-(A+B)")) ~ score_description,
        str_detect(score_description, "\\(A\\+B\\)|\\(C\\+D\\)") ~ str_replace_all(
        score_description,
        "\\(A\\+B\\)|\\(C\\+D\\)",
        "[(C+D)-(A+B)]"
      ),
    TRUE ~ score_description)) %>%
  mutate(
    deck_selection_formula = str_replace_all(
      deck_selection_formula,
      "(late\\(C\\+D\\)-early\\(C\\+D\\))",
      "late[(A+B)-(C+D)]-early[(A+B)-(C+D)]"
    ),
    deck_selection_formula = case_when(
      deck_selection_formula %in% c("(A+B)", "(C+D)") ~ "(C+D)-(A+B)",
      TRUE ~ deck_selection_formula
    ) 
  ) %>%
  mutate(
    score_description       = str_replace_all(score_description,       "\\(B\\+D\\)", "(A+C)"),
    deck_selection_formula  = str_replace_all(deck_selection_formula,  "\\(B\\+D\\)", "(A+C)")
  )

  
# add variable category 
scores_monotonic <- scores_monotonic %>% 
  mutate(category = case_when(
    scoring_approach == "money" ~ "money",
    scoring_approach == "choice behavior" ~ "choice behavior",
    str_detect(scoring_approach, "model") ~ "model parameter",
    scoring_approach == "indeterminable" ~ "indeterminable",
    str_detect(score_description, "categorized") ~ "categorized score", 
    deck_selection_formula %in% c("A", "B", "C", "D") ~ "single deck score",
    deck_selection_formula == "(C+D)-(A+B)" ~ "net score",
    deck_selection_formula == "(A+C)" ~ "loss frequency based score",
    str_detect(deck_selection_formula, "early") ~ "difference score",  
    str_detect(score_description, "one lag") ~ "autocorrelation", 
    monotonic_score_id %in% c(4023, 4013, 2006) ~ "score across two variants", 
    TRUE ~ "other"
  ))

count_scores_monotonic <- nrow(scores_monotonic)

count_scores_monotonic |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Export
Here, we write the table of distinct scores into an Excel file, which is used in scoring_methods.Rmd.
```{r export file}

# select relevant variables
scores_monotonic <- scores_monotonic %>%
  select(monotonic_score_id, score_description, scoring_approach, category, deck_selection_formula, money_formula, choice_behavior_formula, n_trials, subset_trials)

# write into excel file
write_xlsx(scores_monotonic, path = "../../data/processed/meta-methods review/scores_overview.xlsx")

```

# Session Info

```{r session-info}
sessionInfo()
```
