---
title: "Create Scores Overview"
author: "Annika Kuelpmann"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = TRUE)

```

# Preparations 
## Load Packages
```{r packages}
#| message: false
#| warning: false

# load R packages 
library(knitr)
library(tidyverse)
library(readxl)
library(writexl)
library(conflicted)
library(kableExtra)

# declare preferences regarding conflicts 
conflicts_prefer(dplyr::filter)

```

## Load Data 
```{r housekeeping}

# read scoring master file
data_scoring <- read_xlsx("../../data/raw/metamethod review/master coding/codebook_scoring_master.xlsx") 

# read monotonic scores file
# this is a file where we identified scores that can be linear transformed into each other
monotonic <- read_xlsx("../../data/raw/metamethod review/master coding/monotonic_scores.xlsx")

```

# Distinct Scores
This code creates a an object with distinct scores as described in the IGT articles.

```{r overview raw}

# create object with dinstinct scores
scores_overview <- data_scoring %>%
  distinct(score_id, .keep_all=TRUE)

count_scores <- nrow(scores_overview)

count_scores |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Monotonic Scores
Some scores in the sample are described and reported as if they were distinct from each other, but they can be transformed into each other which perfectly preserving ranks. Here, we collapse these scores to prevent an overestimation of the total number of distinct scores. 

```{r overview monotonic}
# join info on monotonic scores to overview of all scores
scores_monotonic <- scores_overview %>% 
  left_join(monotonic, by = "score_id") %>% 
  mutate(monotonic_score_id = ifelse(is.na(new_score_id), score_id, new_score_id)) %>% 
  distinct(monotonic_score_id, .keep_all=TRUE)

# align score descriptions and formulas of monotonic scores 
scores_monotonic <- scores_monotonic %>%
 mutate(
    score_description = case_when(
        str_detect(score_description, fixed("[(C+D)-(A+B)]")) |
        str_detect(score_description, fixed("(C+D)-(A+B)")) ~ score_description,
        str_detect(score_description, "\\(A\\+B\\)|\\(C\\+D\\)") ~ str_replace_all(
        score_description,
        "\\(A\\+B\\)|\\(C\\+D\\)",
        "[(C+D)-(A+B)]"
      ),
    TRUE ~ score_description)) %>%
  mutate(
    deck_selection_formula = str_replace_all(
      deck_selection_formula,
      "(late\\(C\\+D\\)-early\\(C\\+D\\))",
      "late[(A+B)-(C+D)]-early[(A+B)-(C+D)]"
    ),
    deck_selection_formula = case_when(
      deck_selection_formula %in% c("(A+B)", "(C+D)") ~ "(C+D)-(A+B)",
      TRUE ~ deck_selection_formula
    ) 
  ) %>%
  mutate(
    score_description       = str_replace_all(score_description,       "\\(B\\+D\\)", "(A+C)"),
    deck_selection_formula  = str_replace_all(deck_selection_formula,  "\\(B\\+D\\)", "(A+C)")
  )

  
# add variable category 
scores_monotonic <- scores_monotonic %>% 
  mutate(category = case_when(
    scoring_approach == "money" ~ "money",
    scoring_approach == "choice behavior" ~ "choice behavior",
    str_detect(scoring_approach, "model") ~ "model parameter",
    scoring_approach == "indeterminable" ~ "indeterminable",
    str_detect(score_description, "categorized") ~ "categorized score", 
    deck_selection_formula %in% c("A", "B", "C", "D") ~ "single deck score",
    deck_selection_formula == "(C+D)-(A+B)" ~ "net score",
    deck_selection_formula == "(A+C)" ~ "loss frequency based score",
    str_detect(deck_selection_formula, "early") ~ "difference score",  
    str_detect(score_description, "one lag") ~ "autocorrelation", 
    monotonic_score_id %in% c(4023, 4013, 2006) ~ "score across two variants", 
    TRUE ~ "other"
  ))

# add variable: score implemented in graphic table 
scores_monotonic <- scores_monotonic %>% 
    mutate(implemented = case_when(
    monotonic_score_id %in% c(3005, 3025) ~ "yes",
    monotonic_score_id %in% c(4013, 4016, 4023) ~ "too vague to implement",
    monotonic_score_id >= 3001 & monotonic_score_id <= 3042 ~ "too vague to implement",
    monotonic_score_id >= 4039 & monotonic_score_id <= 4054 ~ "referred to in table notes",
    TRUE ~ "yes"
  ))

count_scores_monotonic <- nrow(scores_monotonic)

count_scores_monotonic |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Print 
Here, we print the table of score into an excel file.
```{r print file}

# select relevant variables
scores_monotonic <- scores_monotonic %>%
  select(monotonic_score_id, score_description, scoring_approach, category, deck_selection_formula, money_formula, choice_behavior_formula, n_trials, subset_trials)

# write into excel file
write_xlsx(scores_monotonic, path = "../../data/processed/meta-methods review/scores_overview.xlsx")

```