---
title: "Analyses of IGT Procedures"
author: "Annika Kuelpmann"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = TRUE)

```


```{r packages}
#| include: false

# load R packages 
library(knitr)
library(tidyverse)
library(readxl)
library(conflicted)

# declare preferences regarding conflicts 
conflicts_prefer(dplyr::filter)

```


```{r housekeeping}
#| include: false

# read procedure file
data <- read_xlsx("../../data/raw/metamethod review/master coding/codebook_procedure_master.xlsx") 

```


```{r functions}
#| include: false

# Defining functions used in this script

# function 1: overview over a variable
overview_variable <- function(data, variable) {
  data %>%
    select(all_of(variable)) %>%
    group_by(across(all_of(variable))) %>%
    summarise(n = n(), .groups = "drop") %>%
    mutate(percentage = round(n / sum(n) * 100, 2)) %>%
    arrange(desc(n))
}


# function 2: count and calculate percentages for certain values of a subset of variables
value_counts <- function(data, value_to_count, columns = NULL) {
  if (is.null(columns)) {
    columns <- colnames(data)
  } else {
    columns <- intersect(columns, colnames(data))
  }
  results <- list()
  for (col in columns) {
    count <- sum(data[[col]] == value_to_count, na.rm = TRUE)
    percentage <- round((count / nrow(data)) * 100, 2)
    results[[col]] <- c("count" = count, "percentage" = percentage)
  }
  summary_table <- do.call(rbind, results) %>%
    as.data.frame(stringsAsFactors = FALSE) %>%
    tibble::rownames_to_column("Variable")
  return(summary_table)
}

```


```{r data preparation}
#| include: false

### Task ID 
# add id for each task that was coded
data <- data %>%
  mutate(task_id = sprintf("%03d", row_number())) 

```

## IGT Variants

This section provides an overview of the IGT variants employed in the sample. 
```{r}
#| echo: false

### Data Cleaning
# the variable "protocol" is difficult to read, therefore cleaning it up here 
# Define a mapping from old names to new names
protocol_renaming <- c(
  "indeterminable" = "indeterminable",
  "new protocol" = "new variant",
  "Bechara et al. (1994)" = "original task",
  "Bechara, 2007" = "clinical variant",
  "Bechara, A., Damasio, H., & Damasio, A. R. (2000). Emotion, decision-making, and the orbitofrontal cortex. Cerebral Cortex, 10(3), 295–307." = "Bechara, Damasio, & Damasio, 2000",
    "protocol described in Bechara, Damasio & Damasio 2000 (ABCD)" = "Bechara, Damasio & Damasio, 2000 (ABCD)",
  "Bechara, Tranel & Damasio 2000 (E'F'G'H')" = "Bechara, Tranel & Damasio, 2000 (E'F'G'H')",
  "Cauffman et al., 2010" = "Cauffman, 2010",
  "van den Bos et al., 2006" = "van den Bos, 2006",
  "Crone EA, Somsen RJ, Van BB, et al. Heart rate and skin conductance analysis of antecendents and consequences of decision making. Psychophysiology 2004;41:53140" = "Crone, 2004",
  "Mueller, S. (2012). The psychology experiment building language (Version 0.13.)" = "Mueller, 2012",
  "Becker MP, Collins PF, & Luciana M (2014). Neurocognition in college-aged daily marijuana users. Journal of Clinical and Experimental Neuropsychology, 36(4), 379–98" = "Becker, 2014",
  "protocol described in Windmann 2006" = "Windmann, 2006",
  "Overman et al. (2004)" = "Overman, 2004",
  "Mintzer, M. Z., & Stitzer, M. L. (2002). Cognitive impairment in methadone maintenance patients." = "Mintzer & Stitzer, 2002",
  "Quednow BB, Kuhn KU, Hoppe C, Westheide J, Maier W, Daum I, Wagner M (2007) Elevated impulsivity and impaired decision-making cognition in heavy users of MDMA (‘‘Ecstasy’’)." = "Quednow, 2007",
  "Christakou, Brammer, et al. (2009)" = "Christakou, 2009"
)

# Rename the protocol values
data <- data %>%
  mutate(protocol = recode(protocol, !!!protocol_renaming))

# get an overview over task variants 
variants <- overview_variable(data, "protocol")
print(variants)

```

In order to organise the variants in a systematic manner, they have been aggregated into the following categories:
```{r}
#| echo: false

# aggregate task variants
data <- data %>%
  group_by(task_id) %>%
  mutate(protocol_aggr = case_when(
    protocol %in% c("original task", "clinical variant", "new variant", "indeterminable") ~ protocol,
    TRUE ~ "other cited variant"
  )) %>%
  ungroup()

# get an overview over aggregated task variants
variants_aggr <- overview_variable(data, "protocol_aggr")
print(variants_aggr)

```

## Modifications 

We conducted a descriptive analysis of the modifications that were made to the IGT with respect to different variables. 

### Number of Trials

The frequency of alterations to the number of trials was examined. The variable "n" refers to the number of tasks with the corresponding number of trials, represented as "n_trials". 
```{r}
#| echo: false

# get an overview over number of trials  
trials_n <- overview_variable(data, "n_trials")
print(trials_n)

# aggregate numbers of trials
data <- data %>%
  group_by(task_id) %>%
  mutate(trials_aggr = case_when(
    n_trials %in% c("100", "indeterminable") ~ n_trials,
    TRUE ~ "deviating"
  )) %>%
  ungroup()

# get an overview over aggregated number of trials  
trials_aggr <- overview_variable(data, "trials_aggr")
print(trials_aggr)

# range of number of trials
range_n_trials <- data %>%
  filter(n_trials != "indeterminable") %>%
  summarise(
    min_trials = min(as.numeric(n_trials), na.rm = TRUE),
    max_trials = max(as.numeric(n_trials), na.rm = TRUE),
    valid_observations = n()
  )
print(range_n_trials)

```

### Instructions and Feedback

A descriptive analysis was conducted to determine whether the instructions, the primary goal of the task which can be considered a part of the instruction, and the feedback provided to the participants were consistent with the original task or modified. Additionally, we aimed to quantify the number of instances where these elements could not be determined (value "indeterminable"). The variable "n" represents the number of tasks for which this is the case. 
```{r}
#| echo: false

# get an overview over IGT goal as a part of instruction  
goal <- overview_variable(data, "goal")
print(goal)

# get an overview over information on instructions
instructions <- overview_variable(data, "instructions")
print(instructions)

# get an overview over information on feedback
feedback <- overview_variable(data, "feedback")
print(feedback)

```

Here, the variables of goal, instructions, and feedback were combined into a single variable, named "characteristics". 
```{r}
#| echo: false

# add summary variable for general task characteristics (goal, instructions, feedback)
data <- data %>%
  group_by(task_id) %>%
  mutate(characteristics = case_when(
    all(protocol == "Bechara, 1994") ~ "standard characteristics",
    all(c(goal, instructions, feedback) == "indeterminable") ~ "completely indeterminable characteristics",
    all(grepl("deviating", c(goal, instructions, feedback))) ~ "deviating characteristics",
    any(grepl("deviating", goal) | grepl("deviating", instructions) | grepl("deviating", feedback)) ~ "partly deviating characteristics",
    any(goal == "indeterminable" | instructions == "indeterminable" | feedback == "indeterminable") ~ "indeterminable characteristics",
    TRUE ~ "standard characteristics"
  )) %>%
  ungroup()

# get an overview over information on general task characteristics
characteristics <- overview_variable(data, "characteristics")
print(characteristics)

```

### Loans and Currencies

Here, we present the currencies and capitals that were employed in our sample.
```{r}
#| echo: false

# get an overview over currencies 
currency <- overview_variable(data, "currency")
print(currency)

# get an overview over loans
loan <- overview_variable(data, "starting_capital")
print(loan)

# create data frame with combination of loans and currencies
capital <- data %>%
  select(task_id, starting_capital, currency) %>%
  group_by(starting_capital, currency) %>%
  summarise(n = n(), .groups = 'drop') %>%
  mutate(percentage = round(n / sum(n) * 100, 2)) %>%
  arrange(currency)
print(capital)

```

### Gain Loss Schedules

A number of variables were examined which reflect the gain-loss schedule of the tasks in use, including the fixed gains and losses for each deck and the frequencies of gains and losses for a given interval of trials. These variables were then aggregated to create a new variable which indicates whether the gain-loss schedule is similar to that of the original task, or whether it is deviating from it, and/or whether it contains elements which are indeterminable. 

```{r}
#| echo: false

# add aggregated variable for gain loss schedule information to data
data <- data %>%
  group_by(task_id) %>%
  mutate(schedule = case_when(
    all(protocol == "Bechara, 1994") ~ "standard schedule",
    all(
      c(
        reward_adv == 50, reward_disadv == 100, 
        gain1 == 250, gain2 == 250, loss1 == -250, loss2 == -250, 
        gain_int1 == 10, gain_int2 == 10, loss_int1 == 10, loss_int2 == 10,
        freq_adv1 == 0.5, freq_disadv1 == 0.5, freq_adv2 == 0.1, freq_disadv2 == 0.1
      )
    ) ~ "standard schedule",
    all(
      c(
        reward_adv, reward_disadv, 
        gain1, gain2, loss1, loss2, 
        gain_int1, gain_int2, loss_int1, loss_int2,
        freq_adv1, freq_adv2, freq_disadv1, freq_disadv2
      ) == "indeterminable"
    ) ~ "completely indeterminable schedule", 
    any(
      c(
        reward_adv, reward_disadv, 
        gain1, gain2, loss1, loss2, 
        gain_int1, gain_int2, loss_int1, loss_int2,
        freq_adv1, freq_adv2, freq_disadv1, freq_disadv2
      ) == "indeterminable"
    ) ~ "partly indeterminable schedule", 
    TRUE ~ "deviating schedule"
  )) %>%
  ungroup()

# get an overview over schedules 
schedules <- overview_variable(data, "schedule")
print(schedules)

```

### Incentives

Here, we sought to determine whether the incentives provided to participants were contingent upon their performance in the task or not. 
```{r}
#| echo: false

# get an overview over incentives 
incentives <- overview_variable(data, "incentives")
print(incentives)

```

## Lack of Reporting
One of the objectives of this study was to ascertain the degree of transparency in the reporting of IGT procedures.

### Summary of Indeterminables
In order to achieve this, we demonstrate, for each variable, the frequency of instances where an indeterminable value was assigned. The variable "count" represents the number of tasks for which this occurred. 

```{r}
#| echo: false

# create data frame with relevant variables
data_ind <- data %>%
  select(goal, instructions, feedback, currency, starting_capital, 
         contains("reward"), contains("1"), contains("2"), 
         contains("aggr"))

# count how often variable is set to indeterminable
indeterminables <- value_counts(data_ind, "indeterminable") %>%
  arrange(desc(count))

# print data frame with frequency of indeterminable values 
print(indeterminables)

```