---
title: "IGT Scoring Methods"
author: "Annika Kuelpmann"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = TRUE)

```

```{r packages}
#| include: false

# load R packages 
library(knitr)
library(tidyverse)
library(readxl)
library(conflicted)
library(psych)
library(viridis)
library(viridisLite)
library(scales)
library(forcats)
library(kableExtra)

# declare preferences regarding conflicts 
conflicts_prefer(dplyr::filter)

```

```{r housekeeping}
#| include: false

# read scoring master file
data <- read_xlsx("../../data/raw/metamethod review/master coding/codebook_scoring_master.xlsx") 
# read scores overview file
# this is a file with one row per distinct score 
scores_overview <- read_xlsx("../../data/raw/metamethod review/master coding/all_scores_overview.xlsx")
# read monotonic scores file
# this is a file where we identified scores that can be linear transformed into each other
monotonic_scores <- read_xlsx("../../data/raw/metamethod review/master coding/monotonic_scores.xlsx")

```

```{r functions}
#| include: false

# Defining functions used in this script

# function 1: overview of a variable
overview_variable <- function(data, variable) {
  data %>%
    select(all_of(variable)) %>%
    group_by(across(all_of(variable))) %>%
    summarise(frequency = n(), .groups = "drop") %>%
    mutate(percentage = round(frequency / sum(frequency) * 100, 2)) %>%
    arrange(desc(frequency))
}

# function 2: summary of frequencies of a variable
freq_summary <- function(data) {
  data %>%
    group_by(frequency) %>%
    summarise(count = n(), .groups = 'drop') %>%
    arrange(desc(frequency))
}

```

```{r data preparation}
#| include: false

# add task_id to data 
# and select relevant variables 
data <- data %>%
  mutate(task_id = group_indices(., article_id, variant)) %>%
  select(article_id, task_id, score_count, score_id,
         n_trials, subset_trials)

# join info on monotonic scores 
# and add variable with respective ID for each score 
data <- data %>% 
  left_join(monotonic_scores, by = "score_id") %>% 
  mutate(monotonic_score_id = ifelse(is.na(new_score_id), score_id, new_score_id)) 

# add frequencies for monotonic scores 
data <- data %>%
  group_by(monotonic_score_id) %>%
  mutate(monotonic_score_id_freq = n()) %>%
  ungroup()

# join info on monotonic scores to overview of all scores
scores_overview <- scores_overview %>% 
  left_join(monotonic_scores, by = "score_id") %>% 
  mutate(monotonic_score_id = ifelse(is.na(new_score_id), score_id, new_score_id)) %>% 
  select(score_id, monotonic_score_id, score_description, scoring_approach, category, n_trials, subset_trials)

```

## Number of Scores

### Total Number of Scores

In our sample of 106 tasks, this is the total number of scores we identified:

```{r number1}

# count the total number of scores across the whole sample
# Count the number of rows in the data
total_of_scores <- nrow(data)

total_of_scores |>
  kable() |>
  kable_classic(full_width = FALSE)


```

### Number of Scores per Task 

We analyzed how many scores were used per task; "number_of_scores" represents the number of scores within a given task, "number_of_tasks" reflects the total number of tasks that had that specific number of scores.

```{r number2}

# count the number of scores per task 
scores_per_task <- overview_variable(data, "task_id") 

# summarize the counts of scores per task 
scores_per_task_summary <- freq_summary(scores_per_task) %>%
  rename(number_of_scores = frequency, number_of_tasks = count)

# print the tibble
scores_per_task_summary |>
  kable() |>
  kable_classic(full_width = FALSE)

```

```{r number2plot}

plot1 <- ggplot(scores_per_task_summary, 
                aes(x = number_of_scores, y = number_of_tasks)) +
  geom_bar(stat = "identity", fill = "#26828E") +
  labs(
    title = "Number of Reported Scores per Task",
    x = "Number of Scores",
    y = "Number of Tasks"
  ) +
  theme_minimal() 

print(plot1)

```

### Summary of Number of Scores per Task

```{r number3}

# summary of number of scores per task 
summ_per_task <- describe(scores_per_task$frequency)

summ_per_task |>
  kable() |>
  kable_classic(full_width = FALSE)

```

The total number and the percentage of tasks with a score range of 1 to 8 are as follows: 

```{r number4}

# proportion of tasks with 1-8 scores
# get the number of tasks with number of scores 8 or less
scores_per_task_minor <- scores_per_task_summary %>% 
  filter(number_of_scores <= 8) %>%
  summarise(sum_tasks = sum(number_of_tasks, na.rm = TRUE)) %>%
  pull(sum_tasks)

scores_per_task_minor  |>
  kable() |>
  kable_classic(full_width = FALSE)

percent_scores_per_task_minor <- scores_per_task_minor / 107 * 100

percent_scores_per_task_minor |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Distinct Scores

This section aims to present a descriptive overview of the distinct scores identified in the sample.

```{r distinct1}

# An overview over the scores that we identified in our sample
scores_overview <- scores_overview %>%
  distinct(score_id, .keep_all = TRUE) %>%
  arrange(score_id)

# scores after collapsing monotonic scores
scores_collapsed <- scores_overview %>%
  distinct(monotonic_score_id, .keep_all = TRUE) %>%
  arrange(monotonic_score_id)

```


### Number of Distinct Scores

This reflects the total number of distinct scores identified in our sample, both prior to and following the collapsing of scores that were initially identified as distinct but were also linear transformations of one another. All of the following analyses are done with the collapsed scores.

```{r distinct2}

# count the number of scores we identified in our sample 
count_scores <- nrow(scores_overview)

count_scores |>
  kable() |>
  kable_classic(full_width = FALSE)


# count the number of scores after collapsing monotonic scores into one score
count_scores_collapsed <- nrow(scores_collapsed)

count_scores_collapsed |>
  kable() |>
  kable_classic(full_width = FALSE)


```

### Reusage of Distinct Scores

This illustrates the frequency with which scores were reused within the sample. The variable "times_reused" reflects the number of instances in which a score was reused within the sample, while "number_of_scores" provides the total number of scores that were reused at that frequency.

```{r dinstinct3}

# Count how many scores are reused and how often they are reused
scores_reusage <- overview_variable(data, "monotonic_score_id") %>%
  freq_summary() %>%
  rename(times_reused = frequency, number_of_scores = count)

scores_reusage |>
  kable() |>
  kable_classic(full_width = FALSE)


```


```{r distinct3plot}

plot2 <- ggplot(scores_reusage, 
                aes(x = number_of_scores, y = times_reused, fill = number_of_scores)) +
  geom_col(stat = "identity", fill = "#31688E") +
  theme_minimal() +
  labs(    
    title = "Number of Scores by Frequency in the Sample",
    x = "Number of Scores",
    y = "Frequency of a Score in the Sample"
    ) +
  theme(legend.position = "none") 

print(plot2)

```

In order to simplify the overview, we have categorized the reuse of scores according to the following criteria: A unique score is defined as a score that has never been reused within the sample. A common score is one that has been reused more than ten times. Finally, a rare score corresponds to a score that has been reused up to nine times. 

```{r dinstinct4}

# Add score_freq_category variable to data 
data <- data %>%
  mutate(score_freq_category = case_when(
    monotonic_score_id_freq == 1 ~ "unique",
    monotonic_score_id_freq < 10 ~ "rare",
    monotonic_score_id_freq >= 10 ~ "common"
  ))

scores_frequencies <- overview_variable(data, "score_freq_category")

scores_frequencies |>
  kable() |>
  kable_classic(full_width = FALSE)


```

### Types of Scores 

The following section presents a quantification of the scores, categorized according to the types identified in the sample. "frequency" reflects the number of distinct scores that have been categorized within that particular group.

```{r types1}

# proportion of scores per type in sample 
score_types_total <- data %>%
  select(task_id, monotonic_score_id) %>%
  left_join(scores_collapsed, by = "monotonic_score_id") %>%
  overview_variable("category") 

# proportion of distinct scores per type 
score_types <- overview_variable(scores_collapsed, "category") 

score_types |>
  kable() |>
  kable_classic(full_width = FALSE)


```

A closer look at the tasks that used model parameters: 

```{r types2}

# Count the occurrences of model parameters across tasks 
model_parameters <- data %>%
  select(task_id, monotonic_score_id) %>%
  left_join(scores_collapsed, by = "monotonic_score_id") %>%
  filter(category == "model parameter")

# view the tasks that used igt models as scoring approaches 
igt_models <- model_parameters %>%
   select(task_id, scoring_approach) %>%
   distinct(task_id, .keep_all = TRUE)

igt_models |>
  kable() |>
  kable_classic(full_width = FALSE)

```


## Common Scores

This is an analysis of the most frequently used scores in our sample. The 10 most prevalent scores were the net score across all trials (1001), the net scores across five blocks of 20 trials (1002, 1003, 1004, 1005, 1006), and the deck selections across all trials for each of the four decks A to D (1011, 1020, 1029, 1038). These scores were reused more than ten times across the whole sample and thus are matching the "common" category introduced above. 

```{r common1}

# find the most frequently used scores
# Count the occurrences of each score_id
score_counts <- overview_variable(data, "monotonic_score_id") 

# identify the 10 most common scores
top_10_scores <- score_counts %>%
  top_n(10, wt = frequency) 

top_10_scores |>
  kable() |>
  kable_classic(full_width = FALSE)

```

The proportion of these 10 common scores within our sample is as follows:

```{r common2}

### find out how large the proportion of the most frequent scores is 
# filter the data frame to keep only the rows with the top 10 scores
top_10_scores <- top_10_scores %>%
  pull(monotonic_score_id)

filtered_data_top_10 <- data %>%
  filter(monotonic_score_id %in% top_10_scores)

# Count the occurrences of these top 10 scores
top_10_score_counts <- filtered_data_top_10 %>%
  summarise(count = n())

# Calculate the proportion of the top 10 scores
proportion_top_10 <- (top_10_score_counts$count / total_of_scores) * 100

proportion_top_10 |>
  kable() |>
  kable_classic(full_width = FALSE)


```

This is the number of tasks that would use at least one of the 10 common scores, as well as the number of tasks that would use only common scores:

```{r common3}

# find out how many tasks use at least one of them
# Count the number of unique task_ids associated with these top 10 scores
tasks_with_top_10 <- filtered_data_top_10 %>%
  select(task_id) %>%
  distinct() 

number_tasks_with_top_10 <- nrow(tasks_with_top_10)

number_tasks_with_top_10 |>
  kable() |>
  kable_classic(full_width = FALSE)


# Count the number of unique task_ids associated with only top 10 scores (reporting no other scores)
tasks_with_only_top_10 <- data %>%
  group_by(task_id) %>%
  filter(all(monotonic_score_id %in% top_10_scores)) %>%
  distinct(task_id)

number_tasks_with_only_top_10 <- nrow(tasks_with_only_top_10)

number_tasks_with_only_top_10 |>
  kable() |>
  kable_classic(full_width = FALSE)

```

```{r common3plot}

# cumulative proportion 
scores_cumulative <- data %>%
  count(monotonic_score_id) %>%  
  arrange(desc(n)) %>%  
  mutate(
    cumulative_count = cumsum(n),  
    cumulative_proportion = cumulative_count / sum(n)  
  )

# Calculate the number of scores at each quantile
quantile_counts <- round(quantile(seq_len(nrow(scores_cumulative)), c(0.25, 0.50, 0.75, 1.00)))

# Calculate positions for quantiles
quantile_positions <- quantile(seq_len(nrow(scores_cumulative)), c(0.25, 0.50, 0.75, 1.00))
# define labels 
quantile_labels <- paste0(
  c("25%", "50%", "75%", "100%"), 
  " (n = ", quantile_counts, ")"
)

# plot 
plot3 <- ggplot(scores_cumulative, 
                aes(x = seq_len(nrow(scores_cumulative)), y = cumulative_proportion)) +
  geom_point(color = "#482878") +
  geom_vline(xintercept = quantile_positions, linetype = "dashed", color = "black") +  
  labs(
    x = "Percentage and Number of Scores",
    y = "Cumulative Proportion",
    title = "Cumulative Distribution of Scores"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent) +  # Convert y-axis labels to percentages
  scale_x_continuous(
    breaks = quantile_positions,
    labels = quantile_labels
  )

print(plot3)

```

### Tasks Using the Net Score

Here we present the number of tasks that employ the net score across all trials, irrespective of the number of other scores provided. Additionally, we present the number of tasks that solely provide the net score without including other scores.

```{r net score}

### Identify how often the standard measures net score and learning index are used
# Filter tasks containing net score across all trials (score_id 1001)
tasks_with_net_score <- data %>%
  select(task_id, monotonic_score_id, score_id) %>%
  filter(monotonic_score_id == 1001)

# Count the number of tasks containing net score across all trials
num_tasks_with_net_score <- tasks_with_net_score %>%
  distinct(task_id) %>%
  nrow()

num_tasks_with_net_score |>
  kable() |>
  kable_classic(full_width = FALSE)


# view the tasks which report more than one monotonic variant of the net score across all trials
# Filter for task_id values that appear more than once
tasks_with_net_score_duplicates <- tasks_with_net_score %>%
  group_by(task_id) %>%               
  filter(n() > 1) %>%             
  ungroup() 

# count the number those tasks 
number_tasks_with_net_score_duplicates <- tasks_with_net_score_duplicates %>%
  distinct(task_id) %>%  
  nrow()

# Identify tasks with net score reported only
tasks_with_net_score_only <- data %>%
  group_by(task_id) %>%
  filter(all(monotonic_score_id == 1001)) %>%
  ungroup()

# Count the number of tasks with net score reported only
num_tasks_with_net_score_only <- tasks_with_net_score_only %>%
  distinct(task_id) %>%
  nrow()

num_tasks_with_net_score_only |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Tasks Using the Learning Index 

Here we present the number of tasks that employed net scores across five blocks of 20 trials in conjunction with one another, thereby providing the conventional learning index.Secondly, the number of tasks providing only this learning index without utilising other scores is presented.

```{r learning index}

# Filter tasks containing learning index across 5 blocks of 20 trials (score_id 1002 to 1006)
# Define the vector of monotonic_score_id values
required_scores <- c(1002, 1003, 1004, 1005, 1006)

tasks_with_learning_index <- data %>%
  filter(monotonic_score_id %in% required_scores) %>%
  group_by(task_id) %>%
  summarise(count_learning_index = n_distinct(monotonic_score_id)) %>%
  filter(count_learning_index == length(required_scores)) %>%
  select(task_id) 

# Count the number of articles containing scores learning index 
num_tasks_with_learning_index <- nrow(tasks_with_learning_index)

num_tasks_with_learning_index |>
  kable() |>
  kable_classic(full_width = FALSE)


# Identify tasks with learning index reported only
tasks_with_learning_index_only <- data %>%
  group_by(task_id) %>%
  filter(all(monotonic_score_id %in% required_scores)) %>%
  ungroup()

# Count the number of tasks with learning index reported only
num_tasks_with_learning_index_only <- tasks_with_learning_index_only %>%
  distinct(task_id) %>%
  nrow()

num_tasks_with_learning_index_only |>
  kable() |>
  kable_classic(full_width = FALSE)

```

Here we present the number of tasks that employ both the net score across all trials, and the learning index, irrespective of other scores. Secondly, we present the number of tasks that provide only the net score and learning index.

```{r both1}

# Find tasks that contain both net score and learning index
# Define the vector of monotonic_score_id values for that case
required_scores_both <- c(1001, 1002, 1003, 1004, 1005, 1006)

tasks_with_both <- data %>%
  filter(monotonic_score_id %in% required_scores_both) %>%
  group_by(task_id) %>%
  summarise(count_both = n_distinct(monotonic_score_id)) %>%
  filter(count_both == length(required_scores_both)) %>%
  select(task_id) 

num_tasks_with_both <- nrow(tasks_with_both)

num_tasks_with_both  |>
  kable() |>
  kable_classic(full_width = FALSE)

# Identify tasks with net score and learning index reported only
tasks_with_both_only <- data %>%
  group_by(task_id) %>%
  filter(all(monotonic_score_id %in% required_scores_both) & 
         all(required_scores_both %in% monotonic_score_id)) %>%
  ungroup()

# Count the number of tasks with net score and learning index reported only
num_tasks_with_both_only <- tasks_with_both_only %>%
  distinct(task_id) %>%
  nrow()

num_tasks_with_both_only  |>
  kable() |>
  kable_classic(full_width = FALSE)

```

Here we present the number of tasks that used either the learning index or the net score across all trials, or both.

```{r both2}

# Find tasks that contain either net score or learning index
tasks_with_either <- tasks_with_net_score %>%
  full_join(tasks_with_learning_index, by = "task_id") %>%
  distinct(task_id)

# Count the number of tasks containing either net score or learning index
num_tasks_with_either <- nrow(tasks_with_either)

num_tasks_with_either  |>
  kable() |>
  kable_classic(full_width = FALSE)

```

### Tasks Using Single Deck Selections

Here we present the number of tasks that employed the single deck selections for decks A, B, C, and D across all trials in conjunction with one another. Secondly, the number of tasks providing only these four scores without utilizing other scores is presented.

```{r both3}

# Filter tasks containing the deck selections for each of the four decks across all trials 
# Define the vector of monotonic_score_id values
required_scores_ABCD <- c(1011, 1020, 1029, 1038)

tasks_with_ABCD <- data %>%
  filter(monotonic_score_id %in% required_scores_ABCD) %>%
  group_by(task_id) %>%
  summarise(count_learning_index = n_distinct(monotonic_score_id)) %>%
  filter(count_learning_index == length(required_scores_ABCD)) %>%
  select(task_id) 

# Count the number of articles containing scores learning index 
num_tasks_with_ABCD <- nrow(tasks_with_ABCD)

num_tasks_with_ABCD  |>
  kable() |>
  kable_classic(full_width = FALSE)

# Identify tasks with learning index reported only
tasks_with_ABCD_only <- data %>%
  group_by(task_id) %>%
  filter(all(monotonic_score_id %in% required_scores_ABCD)) %>%
  ungroup()

# Count the number of tasks with learning index reported only
num_tasks_with_ABCD_only <- tasks_with_ABCD_only %>%
  distinct(task_id) %>%
  nrow()

num_tasks_with_ABCD_only  |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Unique Scores

Here we provide a descriptive analysis of the number of unique scores, defined as those that were present only once in the sample.

```{r unique1}

# count how many scores are unique in the sample
number_of_unique_scores <- scores_reusage %>%
  filter(times_reused == 1)

number_of_unique_scores  |>
  kable() |>
  kable_classic(full_width = FALSE)

```

This is the proportion of unique scores within our sample:

```{r unique2}

### find out how large the proportion of the unique scores is 
# filter the data frame to keep only the rows with unique scores
unique_scores <- score_counts %>%
  filter(frequency == 1) %>%
  pull(monotonic_score_id)

filtered_data_unique_scores <- data %>%
  filter(monotonic_score_id %in% unique_scores)

# Count the occurrences of unique scores
unique_score_counts <- filtered_data_unique_scores %>%
  summarise(count = n())

# Calculate the proportion of the unique scores
proportion_unique_scores <- (unique_score_counts$count / total_of_scores) * 100

proportion_unique_scores |>
  kable() |>
  kable_classic(full_width = FALSE)

```

The following table presents the number of unique scores per task: "unique_scores" indicates the number of scores that are exclusive to a given task, while "frequency" reflects the number of tasks in which that number of unique scores has been obtained.

```{r unique3}

# count the tasks and articles that have unique scores 
# create data frame
unique_counts <- data %>%
  left_join(score_counts, by = "monotonic_score_id") %>%
  select(task_id, monotonic_score_id, frequency) %>%
  group_by(task_id, monotonic_score_id) %>%
  summarise(unique_scores = any(frequency == "1"), .groups = 'drop')

# count tasks with and without unique scores
unique_counts <- unique_counts %>%
  group_by(task_id) %>%
  summarise(
    uniques = sum(unique_scores == TRUE),
    recurrings = sum(unique_scores == FALSE)
  )

# summarize the frequencies of unique scores within tasks 
uniques_per_task <- overview_variable(unique_counts, "uniques") %>%
  rename(unique_scores = uniques)

uniques_per_task |>
  kable() |>
  kable_classic(full_width = FALSE)

```

```{r unique3plot}
unique_counts <- unique_counts %>%
  mutate(task_id = as.factor(task_id)) %>%
  arrange(desc(uniques)) %>%
  mutate(rank = row_number())

plot4 <- ggplot(unique_counts, aes(x = factor(rank), y = uniques)) +
  geom_bar(stat = "identity", fill = "#35B779") +
  theme_minimal() +
  labs(
    title = "Number of Unique Scores per Task (Ranked)",
    x = "Rank of Task (by Unique Scores)",
    y = "Number of Unique Scores"
  ) +
  theme(
    axis.text.x = element_blank()) +
  coord_cartesian(xlim = c(0, nrow(unique_counts) + 5))

print(plot4)

```

The following table demonstrates the degree of overlap between tasks that have been assigned unique, rare, or common scores.

```{r unique4}
#| echo: false

# Create condition counts for each task_id
condition_counts <- data %>%
  group_by(task_id) %>%
  summarise(
    has_common = any(score_freq_category == "common"),
    has_rare = any(score_freq_category == "rare"),
    has_unique = any(score_freq_category == "unique"),
    .groups = 'drop'
    ) %>%
  # Count the occurrences of each combination
  group_by(has_common, has_rare, has_unique) %>%
  summarise(count = n(),
            .groups = 'drop') 

# View the resulting condition_counts dataframe
condition_counts |>
  kable() |>
  kable_classic(full_width = FALSE)

```

## Indeterminable Scores

An indeterminable score is a score with elements that are not clearly defined or that has not been adequately described. We have classified indeterminable scores as a unique, given that it was not possible to determine whether an indeterminable score was distinct or not. The total number of indeterminable scores identified in our sample is as follows: 

```{r indeterminable1}

indeterminable_scores <- scores_overview %>%
  filter(category == "indeterminable")

number_of_indeterminable_scores <- nrow(indeterminable_scores)

number_of_indeterminable_scores |>
  kable() |>
  kable_classic(full_width = FALSE)


```

Here we present the number of indeterminate scores per task. "indeterminable_scores" represents the number of scores identified as indeterminable within a given task, "frequency" reflects the number of tasks with that specific number of indeterminable scores.

```{r indeterminable2}

# count how many articles have with indeterminable scores
# create data frame
indeterminable_counts <- data %>%
  left_join(scores_overview, by = "score_id") %>%
  select(article_id, task_id, score_id, category) %>%
  group_by(article_id, task_id, score_id) %>%
  summarise(indeterminable = any(category == "indeterminable"),
            .groups = 'drop')

# overview over articles with and without indeterminable scores
indeterminable_counts <- indeterminable_counts %>%
  group_by(article_id, task_id) %>%
  summarise(
    indeterminable_scores = sum(indeterminable == TRUE),
    determinable_scores = sum(indeterminable == FALSE),
    .groups = 'drop'
  )

# summarize number of indeterminable scores 
indeterminable_counts_summary <- overview_variable(indeterminable_counts, "indeterminable_scores")

indeterminable_counts_summary |>
  kable() |>
  kable_classic(full_width = FALSE)

```
