---
title: "IGT multiverse"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# TODO

- correct wins/losses scoring methods for number of trials in the task


```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

# Get data

```{r}

# dependencies
library(tidyverse)
library(knitr)
library(kableExtra)
#library(sjPlot)
#library(forcats)
library(janitor)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, janitor::round_half_up, digits = n_digits)
}

# get data
data_processed_participant_level <- read_csv("../../data/processed/multiverse/data_processed_participant_level.csv")

```

# Correlate scores

```{r}

# count scoring strategies
n_scoring_strategies <- data_processed_participant_level |>
  select(-id) |>
  colnames() |>
  length()

paste("Number of scoring strategies:", n_scoring_strategies)

# Correlation table
cors_decks_combined <- data_processed_participant_level |>
  select(-id) |>
  relocate(C_plus_D_minus_A_plus_B, .after = "proportion_switch_to_cd_after_loss_on_ab") |>
  cor() |>
  round(2)

diag(cors_decks_combined) <- NA

```

## Between all scoring methods

```{r}

# Correlation between all scoring methods
cors_decks_combined[lower.tri(cors_decks_combined, diag = TRUE)] <- NA

data_cors_combined <- tibble(cors = cors_decks_combined[upper.tri(cors_decks_combined)])

## native 
ggplot(data_cors_combined, aes(cors)) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Correlation between all scoring methods") +
  xlab(expression(paste("Pearson's ",italic("r")))) +
  coord_cartesian(xlim = c(-1,1)) +
  theme_bw()

## absolute
ggplot(data_cors_combined, aes(cors)) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Absolute correlation between all scoring methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

```

## Between modal scoring strategy and all others 

```{r}

# Correlation between modal scoring method and all others
data_cors_combined_modal <- 
  tibble(cors = as_tibble(cors_decks_combined)$C_plus_D_minus_A_plus_B) |>
  drop_na()

## native 
ggplot(data_cors_combined_modal, aes(cors)) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Correlation between modal scoring method and all other methods") +
  xlab(expression(paste("Pearson's ",italic("r")))) +
  coord_cartesian(xlim = c(-1,1)) +
  theme_bw()

## absolute
ggplot(data_cors_combined_modal, aes(abs(cors))) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Absolute correlation between modal scoring method and all other methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

```

## Between top 13 most common scoring methods

```{r}

cors_decks_subset <- data_processed_participant_level |>
  select(C_plus_D_minus_A_plus_B_1, C_plus_D_minus_A_plus_B_2, C_plus_D_minus_A_plus_B_3, C_plus_D_minus_A_plus_B_4, C_plus_D_minus_A_plus_B_5,
         C_plus_D_minus_A_plus_B_block_5_minus_block_1, # not directly a scoring strategy but the use of the most common strategies in the analyses which look at learning/change
         C_plus_D,
         A_plus_B,
         A,
         B,
         C,
         D,
         C_plus_D_minus_A_plus_B) |>
  cor() |>
  round(2)

diag(cors_decks_subset) <- NA

cors_decks_subset

# Correlation between all scoring methods
cors_decks_subset[lower.tri(cors_decks_subset, diag = TRUE)] <- NA

data_cors_subset <- tibble(cors = cors_decks_subset[upper.tri(cors_decks_subset)])

## native 
ggplot(data_cors_subset, aes(cors)) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Correlation between 13 most common scoring methods") +
  xlab(expression(paste("Pearson's ",italic("r")))) +
  coord_cartesian(xlim = c(-1,1)) +
  theme_bw()

## absolute
ggplot(data_cors_subset, aes(cors)) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Absolute correlation between 13 most common scoring methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

```

## Between modal scoring strategy and most common ones 

```{r}

# Correlation between modal scoring method and all others
data_cors_subset_modal <- 
  tibble(cors = as_tibble(cors_decks_subset)$C_plus_D_minus_A_plus_B) |>
  drop_na()

## native 
ggplot(data_cors_subset_modal, aes(cors)) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Correlation between modal scoring method and the next 12 most common methods") +
  xlab(expression(paste("Pearson's ",italic("r")))) +
  coord_cartesian(xlim = c(-1,1)) +
  theme_bw()

## absolute
ggplot(data_cors_subset_modal, aes(abs(cors))) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Absolute correlation between modal scoring method and the next 12 most common methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

```

# EGA

```{r}

# library(EGAnet)
# 
# dat <- data_scored_deck_choices_combined |>
#   rename(A = A, 
#          B = C, 
#          C = C_plus_A, 
#          D = C_minus_A, 
#          E = C_plus_A_late, 
#          F = A_late, 
#          G = C_minus_A_late, 
#          H = C_late, 
#          
#          I = C_minus_A_early, 
#          J = A_early, 
#          K = C_early, 
#          L = C_plus_A_early, 
#          
#          M = D_minus_B_early, 
#          N = D_early, 
#          O = B_early, 
#          P = D_plus_B_early, 
#          Q = C_plus_D_early, 
#          R = A_plus_B_early, 
#          S = C_plus_D_minus_A_plus_B_early, 
# 
#          T = B, 
#          U = D, 
#          V = D_plus_B, 
#          W = D_minus_B, 
#          X = C_plus_D, 
#          Y = A_plus_B, 
#          Z = C_plus_D_minus_A_plus_B, 
#          A1 = B_late, 
#          B1 = D_late, 
#          C1 = D_plus_B_late,
#          D1 = D_minus_B_late,
#          E1 = C_plus_D_late,
#          F1 = A_plus_B_late,
#          G1 = C_plus_D_minus_A_plus_B_late)
# 
# fit <- EGA(data = dat, cor = "pearson")
# 
# 
# 
# dat <- data_scored_deck_choices_combined |>
#   select(-id, -contains("early"), -contains("late")) 
# 
# fit <- EGA(data = dat, cor = "pearson")

```

# simulated criterion

## Max correlation between uncorrelated data at a given N

```{r}

# n <- 10
# max(replicate(n = 1000, abs(cor(rnorm(n), rnorm(n)))))

ns = seq(from = 10, to = 1010, by = 50)
res <- round(sapply(ns, function(n){max(replicate(n = 10000, abs(cor(rnorm(n), rnorm(n)))))}), 2)
names(res) <- paste0("N=", ns)
res

# not working correctly i think:
# df <- data.frame(ns = seq(from = 10, to = 300, by = 10)) |>
#   rowwise() |>
#   mutate(max_cor = map_dbl(ns, function(n) {
#     max(replicate(n = 1000, abs(cor(rnorm(n), rnorm(n)))))
#   }))

```

Population r = 0.0 (random data)

Using subset of 40 random participants.

```{r}

set.seed(42)

data_simulated_criterion <- data_processed_participant_level |>
  mutate(criterion = rnorm(n = nrow(data_processed_participant_level), mean = 0, sd = 1)) |>
  sample_n(size = 40)

```

## Between top 13 most common scoring methods and simulated criterion

```{r}

cors_decks_subset_simulated_criterion <- data_simulated_criterion |>
  select(C_plus_D_minus_A_plus_B_1, C_plus_D_minus_A_plus_B_2, C_plus_D_minus_A_plus_B_3, C_plus_D_minus_A_plus_B_4, C_plus_D_minus_A_plus_B_5,
         C_plus_D_minus_A_plus_B_block_5_minus_block_1, # not directly a scoring strategy but the use of the most common strategies in the analyses which look at learning/change
         C_plus_D,
         A_plus_B,
         A,
         B,
         C,
         D,
         C_plus_D_minus_A_plus_B,
         criterion) |>
  cor() |>
  round(2)

cors_decks_subset_simulated_criterion[lower.tri(cors_decks_subset_simulated_criterion, diag = TRUE)] <- NA

diag(cors_decks_subset_simulated_criterion) <- NA

cors_decks_subset_simulated_criterion

# Correlation between modal scoring method and all others
data_cors_subset_simulated_criterion <- 
  tibble(cors = as_tibble(cors_decks_subset_simulated_criterion)$criterion) |>
  drop_na()

# ## native 
# ggplot(data_cors_subset_simulated_criterion, aes(cors)) +
#   geom_histogram(binwidth = .1, boundary = 0) +
#   ggtitle("Correlation between simulated criterion method and the 13 most common scoring methods") +
#   xlab(expression(paste("Pearson's ",italic("r")))) +
#   coord_cartesian(xlim = c(-1,1)) +
#   theme_bw()

## absolute
ggplot(data_cors_subset_simulated_criterion, aes(abs(cors))) +
  geom_histogram(binwidth = .1, boundary = 0) +
  ggtitle("Absolute correlation between modal scoring method and the 13 most common scoring methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

```


