---
title: "IGT multiverse"
subtitle: "Analysis"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

# Get data

```{r}

# dependencies
library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)
library(patchwork)
library(corrr)

# get data
data_processed_participant_level <- read_csv("../../data/processed/multiverse/data_processed_participant_level.csv") |>
  # remove id
  select(-id) 

```

# Summary stats

Number of unique scoring methods implemented = `r ncol(data_processed_participant_level)`.

Number of participants = `r nrow(data_processed_participant_level)`.

# Correlate scores

```{r}

cors <- data_processed_participant_level |>
  # remove scores with no variance
  select(-A_60_cutoff) |>
  correlate() |> 
  stretch() |> 
  filter(y != x) |> # remove diagonal
  arrange(desc(r))

```

## Between all scoring methods

Note that no correlations are exactly 1 or -1 as these would represent linear transformations of variables that are not meaningfully different. Numerous cases of this were removed when writing the data processing, as we discovered some scoring strategies were in fact linear transformations of one another. That is, scores that correlate 1.0 do exist in the literature, but we have removed them here on the basis of being redundant with other labels for the same (linear transformed) scoring strategy.

```{r}

cors |>
  filter(x < y) |> # keeps only upper triangle by comparing characters alphabetically for comparison
  summarize(n_correlations = n(),
            min_r = min(r),
            max_r = max(r),
            median_r = median(r),
            mad_r = mad(r),
            min_abs_r = min(abs(r)),
            max_abs_r = max(abs(r)),
            median_abs_r = median(abs(r)),
            mad_abs_r = mad(abs(r))) |>
  mutate_if(is.numeric, round_half_up, digits = 3) |>
  kable() |>
  kable_classic(full_width = FALSE)

## native 
p_native_all <- ggplot(cors |> filter(x < y), aes(r)) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#0072B2") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  ggtitle("Correlations among all scoring strategies") +
  #xlab(expression(paste("Pearson's ",italic("r")))) +
  xlab("Correlation") +
  coord_cartesian(xlim = c(-1,1)) +
  theme_bw()

## absolute
p_absolute_all <- ggplot(cors |> filter(x < y), aes(abs(r))) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#0072B2") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  #ggtitle("Absolute correlation between modal scoring strategy and all others") +
  #xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  xlab("Absolute correlation") +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

p_native_all
p_absolute_all

```

## Between modal scoring strategy and all others 

```{r}

cors |>
  filter(x == "C_plus_D_minus_A_plus_B") |> # keeps only upper triangle by comparing characters alphabetically for comparison
  summarize(n_correlations = n(),
            min_r = min(r),
            max_r = max(r),
            median_r = median(r),
            mad_r = mad(r),
            min_abs_r = min(abs(r)),
            max_abs_r = max(abs(r)),
            median_abs_r = median(abs(r)),
            mad_abs_r = mad(abs(r))) |>
  mutate_if(is.numeric, round_half_up, digits = 3) |>
  kable() |>
  kable_classic(full_width = FALSE)

## native 
p_native_modal <- ggplot(cors |> filter(x == "C_plus_D_minus_A_plus_B"), aes(r)) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#E69F00") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  ggtitle("Correlations with modal scoring strategy") +
  #xlab(expression(paste("Pearson's ",italic("r")))) +
  xlab("Correlation") +
  coord_cartesian(xlim = c(-1,1)) +
  theme_bw()

## absolute
p_absolute_modal <- ggplot(cors |> filter(x == "C_plus_D_minus_A_plus_B"), aes(abs(r))) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#E69F00") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  #ggtitle("Absolute correlation between all scoring strategies") +
  #xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  xlab("Absolute correlation") +
  coord_cartesian(xlim = c(0,1)) +
  theme_bw()

p_native_modal
p_absolute_modal

```

## Combined

```{r fig.height=6, fig.width=8}

p_native_all +
  p_native_modal +
  p_absolute_all +
  p_absolute_modal +
  plot_layout(ncol = 2, nrow = 2)

ggsave("plot.pdf", width = 9, height = 6)
ggsave("plot.png", width = 9, height = 6, dpi = 600)

```

# Session info

```{r}

sessionInfo()

```


