---
title: "IGT multiverse"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

# Get data

```{r}

# dependencies
library(tidyverse)
library(knitr)
library(kableExtra)
#library(sjPlot)
#library(forcats)
library(janitor)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, janitor::round_half_up, digits = n_digits)
}

# get data
load("../../data/raw/multiverse/IGTdata.rdata")

```

# Score items

## Win/loss

```{r}

data_wins <- 
  bind_rows(
    wi_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("100trials", id, sep = "_")),
    wi_95 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("95trials", id, sep = "_")),
    wi_150 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("150trials", id, sep = "_"))
  )

data_losses <- 
  bind_rows(
    lo_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("100trials", id, sep = "_")),
    lo_95 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("95trials", id, sep = "_")),
    lo_150 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("150trials", id, sep = "_"))
  )

data_winloss <- 
  full_join(data_wins, data_losses, by = c("id", "trial")) |>
  # dichotomise - TODO check logic 
  mutate(balance = win + loss, 
         win_binary = ifelse(win > 50, 1, 0),  # TODO is > 50 right here?
         loss_binary = ifelse(loss < 0, 1, 0),
         advantagous = ifelse(balance > 0, 1, 0),
         disadvantagous = ifelse(balance < 0, 1, 0))

# check that no advantagous==1 & disadvantagous==1 exist
data_winloss |>
  count(advantagous, disadvantagous)

```

## Choices

This assumes 1 = A, 2 = B etc. Check against the paper?

```{r}

data_choice <- 
  bind_rows(
    choice_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("100trials", id, sep = "_")),
    choice_95 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("95trials", id, sep = "_")),
    choice_150 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("150trials", id, sep = "_"))
  ) |>
  mutate(choice = as.factor(case_when(choice == 1 ~ "A",
                                      choice == 2 ~ "B",
                                      choice == 3 ~ "C",
                                      choice == 4 ~ "D")))

```

## Combine

```{r}

data_combined <- full_join(data_winloss, data_choice, by = c("id", "trial"))

```

# Score participants

## (Dis)advantageous choices 

TODO MAP THESE ONTO SCORING STRATEGIES, NOT SURE IF NEEDED

### All trials

```{r}

# data_scored_advantage <- data_combined |>
#   group_by(id) |>
#   summarize(mean_advantagous = mean(advantagous),
#             mean_disadvantagous = mean(disadvantagous)) |>
#   ungroup() |>
#   mutate(diff = mean_advantagous - mean_disadvantagous)
# 
# data_scored_advantage |>
#   select(-id) |>
#   cor() |>
#   round(2)
# 
# data_scored_advantage |>
#   pivot_longer(cols = -id) |>
#   ggplot(aes(value)) +
#   geom_density() +
#   facet_wrap(~name)

```

### Early and late trials 

Not observed in literature, but plausible. 

```{r}

# TODO

```

### Difference between early and late trials

Observed in literature, but simple early and late are not. 

```{r}

# TODO

```

## Wins/losses

Ignores starting capital

### All trials

```{r}

data_scored_winloss <- data_combined |>
  group_by(id) |>
  summarize(total_won = sum(win),
            total_lost = sum(loss)) |>
  ungroup() |>
  mutate(total_won_minus_lost = total_won + total_lost)

```

### Early and late trials 

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

### Difference between early and late trials

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

## Deck choice counts

### All trials

```{r}

data_scored_deck_choices <- data_combined |>
  group_by(id) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  mutate(C_plus_A  = C+A,
         C_minus_A = C-A,
         D_plus_B  = D+B,
         D_minus_B = D-B,
         C_plus_D  = C+D,
         A_plus_B  = A+B,
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B,
         C_plus_D_divided_by_A_plus_B = C_plus_D/A_plus_B)

data_combined |>
  group_by(choice) |>
  summarize(mean_win = mean(win),
            mean_loss = mean(loss),
            mean_balance = mean(balance))

```

### Early and late trials

```{r}

data_scored_deck_choices_subset_long <- data_combined |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= 0.4 ~ "early",   # TODO cutoffs could be varied
                            trial_proportion >= 0.4 ~ "late")) |>
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  mutate(C_plus_A  = C+A,
         C_minus_A = C-A,
         D_plus_B  = D+B,
         D_minus_B = D-B,
         C_plus_D  = C+D,
         A_plus_B  = A+B,
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B) 

data_scored_deck_choices_subset <- data_scored_deck_choices_subset_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset))

```

### Difference between early and late trials

Inherits the proportion values used to create the early and late subsets from the early/late chunk above. 

```{r}

data_scored_deck_choices_subset_diff <- data_scored_deck_choices_subset_long |>
  pivot_longer(names_to = "variable",
               values_to = "score",
               cols = c(-id, -subset)) |>
  pivot_wider(names_from = subset, 
              values_from = score) |>
  mutate(diff = late - early) |>
  select(-early, -late) |>
  pivot_wider(names_from = variable,
              values_from = diff,
              names_prefix = "diff_")

```

## Deck switching

### All trials

```{r}

# "trials participants switch decks"
data_scored_deck_switching_any <- data_combined |>
  group_by(id) |>
  mutate(switch_any = choice == lag(choice)) |>
  summarize(proportion_switch_any = mean(switch_any, na.rm = TRUE))

# "deck switching after loss"
# proportion of switching vs non switching after loss, excluding trial on which there were not losses. 
# TODO This assumption of excluding non-loss trials needs checking.
data_scored_deck_switching_after_loss <- data_combined |>
  filter(loss_binary == TRUE) |>
  group_by(id) |>
  mutate(switch_after_loss = choice == lag(choice)) |>
  summarize(proportion_switch_after_loss = mean(switch_after_loss, na.rm = TRUE))  

# "card selections from deck A which immediately followed a loss on that deck"
# proportion of switching away from deck A after a loss on deck A
data_scored_deck_switching_a_after_loss_on_a <- data_combined |>
  filter(lag(loss_binary) == TRUE & lag(choice == "A")) |>
  group_by(id) |>
  mutate(switch_from_a_after_loss_on_a = choice != "A") |>
  summarize(proportion_switch_from_a_after_loss_on_a = mean(switch_from_a_after_loss_on_a, na.rm = TRUE))

# "trials participants switch from disadvantageous to advantageous decks after loss (total)"
# proportion of switching to deck C or D after a loss on A or B
data_scored_deck_switching_to_cd_after_loss_on_ab <- data_combined |>
  filter(lag(loss_binary) == TRUE & lag(choice %in% c("A", "B"))) |>
  group_by(id) |>
  mutate(switch_to_cd_after_loss_on_ab = choice %in% c("C", "D")) |>
  summarize(proportion_switch_to_cd_after_loss_on_ab = mean(switch_to_cd_after_loss_on_ab, na.rm = TRUE))
       
# combine
data_scored_deck_switching <- data_scored_deck_switching_any |>
  full_join(data_scored_deck_switching_after_loss, by = "id") |>
  full_join(data_scored_deck_switching_a_after_loss_on_a, by = "id") |>
  full_join(data_scored_deck_switching_to_cd_after_loss_on_ab, by = "id")

```

### Early and late trials

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

### Difference between early and late trials

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

## Combine

```{r}

data_scored_combined <- data_scored_winloss |>
  full_join(data_scored_deck_choices,             by = "id") |>
  full_join(data_scored_deck_choices_subset,      by = "id") |>
  full_join(data_scored_deck_choices_subset_diff, by = "id") |>
  full_join(data_scored_deck_switching,           by = "id")

```

# Correlate scores

```{r}

# count scoring strategies
n_scoring_strategies <- data_scored_combined |>
  select(-id, -C, -C_early, -C_late, -diff_C) |>
  colnames() |>
  length()

paste("Number of scoring strategies:", n_scoring_strategies)

# Correlation table
cors_decks_combined <- data_scored_combined |>
  select(-id, -C, -C_early, -C_late, -diff_C) |>
  cor() |>
  round(2)

diag(cors_decks_combined) <- NA

# Correlation between modal scoring method and all others
data_cors_combined_modal <- 
  tibble(cors = as_tibble(cors_decks_combined)$C_plus_D_minus_A_plus_B) |>
  drop_na()

## native 
ggplot(data_cors_combined_modal, aes(cors)) +
  geom_histogram() +
  ggtitle("Correlation between modal scoring method and all other methods") +
  xlab(expression(paste("Pearson's ",italic("r")))) +
  xlim(-1,1) +
  theme_bw()

## absolute
ggplot(data_cors_combined_modal, aes(abs(cors))) +
  geom_histogram() +
  ggtitle("Absolute correlation between modal scoring method and all other methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  xlim(0,1) +
  theme_bw()

# Correlation between all scoring methods
cors_decks_combined[lower.tri(cors_decks_combined, diag = TRUE)] <- NA

data_cors_combined <- tibble(cors = cors_decks_combined[upper.tri(cors_decks_combined)])

## native 
ggplot(data_cors_combined, aes(cors)) +
  geom_histogram() +
  ggtitle("Correlation between all scoring methods") +
  xlab(expression(paste("Pearson's ",italic("r")))) +
  xlim(-1,1) +
  theme_bw()

## absolute
ggplot(data_cors_combined, aes(cors)) +
  geom_histogram() +
  ggtitle("Absolute correlation between all scoring methods") +
  xlab(expression(paste("|Pearson's ",italic("r"), "|"))) +
  xlim(0,1) +
  theme_bw()

```

# EGA

```{r}

# library(EGAnet)
# 
# dat <- data_scored_deck_choices_combined |>
#   rename(A = A, 
#          B = C, 
#          C = C_plus_A, 
#          D = C_minus_A, 
#          E = C_plus_A_late, 
#          F = A_late, 
#          G = C_minus_A_late, 
#          H = C_late, 
#          
#          I = C_minus_A_early, 
#          J = A_early, 
#          K = C_early, 
#          L = C_plus_A_early, 
#          
#          M = D_minus_B_early, 
#          N = D_early, 
#          O = B_early, 
#          P = D_plus_B_early, 
#          Q = C_plus_D_early, 
#          R = A_plus_B_early, 
#          S = C_plus_D_minus_A_plus_B_early, 
# 
#          T = B, 
#          U = D, 
#          V = D_plus_B, 
#          W = D_minus_B, 
#          X = C_plus_D, 
#          Y = A_plus_B, 
#          Z = C_plus_D_minus_A_plus_B, 
#          A1 = B_late, 
#          B1 = D_late, 
#          C1 = D_plus_B_late,
#          D1 = D_minus_B_late,
#          E1 = C_plus_D_late,
#          F1 = A_plus_B_late,
#          G1 = C_plus_D_minus_A_plus_B_late)
# 
# fit <- EGA(data = dat, cor = "pearson")
# 
# 
# 
# dat <- data_scored_deck_choices_combined |>
#   select(-id, -contains("early"), -contains("late")) 
# 
# fit <- EGA(data = dat, cor = "pearson")

```


