---
title: "IGT multiverse"
subtitle: "Data processing"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# Overview

This file takes openly available IGT data (Steingroever et al., 2015; see below) and scores it using the scoring methods discovered in the meta-methods review. The separate analyses.Rmd file then looks at the correlations among the different scoring methods. I implement as many of the scoring methods discovered in the review as possible. Not all were possible to implement due to lack of information for some of them.

Data taken from:

Steingroever, H., Fridberg, D. J., Horstmann, A., Kjome, K. L., Kumari, V., Lane, S. D., … Wagenmakers, E.-J. (2015). Data from 617 Healthy Participants Performing the Iowa Gambling Task: A “Many Labs” Collaboration. Journal of Open Psychology Data, 3(1), e5. DOI: http://doi.org/10.5334/jopd.ak

Article URL:

https://openpsychologydata.metajnl.com/articles/10.5334/jopd.ak/

Data URL:

https://osf.io/8t7rm/

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

# Get data

```{r}

options(scipen=999)

# dependencies
library(tidyverse)
library(janitor)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, janitor::round_half_up, digits = n_digits)
}

# get data
load("../../data/raw/multiverse/IGTdata.rdata")

```

# Score items

## Win/loss

```{r}

# only the 100 trial version of the task is used, other versions would need scores standardised by number of trials
data_wins <- wi_100 |>
  as_tibble() |>
  rownames_to_column(var = "id") |>
  pivot_longer(cols = -id,
               names_to = "trial",
               values_to = "win") |>
  mutate(trial = as.numeric(str_remove(trial, "Wins_")),
         id = paste("100trials", id, sep = "_"))

data_losses <- lo_100 |>
  as_tibble() |>
  rownames_to_column(var = "id") |>
  pivot_longer(cols = -id,
               names_to = "trial",
               values_to = "loss") |>
  mutate(trial = as.numeric(str_remove(trial, "Losses_")),
         id = paste("100trials", id, sep = "_"))

data_winloss <- 
  full_join(data_wins, data_losses, by = c("id", "trial")) |>
  mutate(balance_for_trial = win + loss, 
         loss_binary = ifelse(loss < 0, 1, 0))

```

## Choices

This assumes 1 = A, 2 = B, 3 = C, 4 = D. This seems to be correct when comparing against the average amount won per deck.

```{r}

data_choice <- choice_100 |>
  as_tibble() |>
  rownames_to_column(var = "id") |>
  pivot_longer(cols = -id,
               names_to = "trial",
               values_to = "choice") |>
  mutate(trial = as.numeric(str_remove(trial, "Choice_")),
         id = paste("100trials", id, sep = "_")) |>
  mutate(choice = as.factor(case_when(choice == 1 ~ "A",
                                      choice == 2 ~ "B",
                                      choice == 3 ~ "C",
                                      choice == 4 ~ "D")))

```

## Combine

```{r}

data_processed_trial_level <- data_winloss |>
  full_join(data_choice, by = c("id", "trial"))

```

## Compare decks

To check that I have correctly labelled the disadvantageous decks (A and B) and the advantageous ones (C and D)

```{r}

data_processed_trial_level |>
  group_by(choice) |>
  summarize(mean_win = mean(win),
            mean_loss = mean(loss),
            mean_balance = mean(balance_for_trial)) |>
  round_df(0)

```

## Check all participants have 100 trials

```{r}

data_processed_trial_level |>
  count(id) |>
  count(n)

```

## Write to disk

```{r}

write_csv(data_processed_trial_level,
          "../../data/processed/multiverse/data_processed_trial_level.csv")

```

# Score participants

## Scores observed in meta-methods review implemented below

- [(C+D)-(A+B)]
  - 100% of trials
  - each 5% of trials
  - each 17% of trials 
  - each 20% of trials 
  - each 25% of trials 
  - each 33% of trials 
  - first 10% of trials 
  - first 40% of trials
  - first 50% of trials
  - first 60% of trials
  - 11-50% of trials 
  - 51-80% of trials 
  - last 20% of trials 
  - last 25% of trials 
  - last 40% of trials
  - last 50% of trials
  - last 60% of trials
  - diff between first 40% of trials and last 60% of trials
  - diff between first 33% of trials and last 33% of trials
  - diff between first 17% of trials and last 17% of trials 
  - mean of two IGT completions (all trials) [cannot be implemented here]
  - dichotomised to less than 0 (impaired) or more than 0 (not impaired)
  - trichotomised to less than -10 (impaired), more than 10 (not impaired), or otherwise borderline
  - trichotomised to binomial test p-value for each participant using 100% of trials; significant disadvantageous choices, significant advantageous choices, and non-significant.
  - last 60% dichotomised using median split.
  - normalised via t-scores, 100% of trials [linear transformation of [(C+D)-(A+B)], not retained]
  - normalised via t-scores, blocks of 20% of trials [linear transformation of [(C+D)-(A+B)], not retained]
  - normalised via t-scores and dichotomised using t-score = 39, 100% of trials
  - normalised via t-scores and dichotomised using t-score = 39, blocks of 20% of trials
  - normalised via t-scores and dichotomised using t-score = 50, 100% of trials
  - normalised via t-scores and dichotomised using t-score = 50, blocks of 20% of trials
- A+B [perfectly correlated with [(C+D)-(A+B)], no need to implement here]
- C+D 
  - 100% of trials [perfectly correlated with [(C+D)-(A+B)], no need to implement here]
  - 100% of trials and logit transformed 
- [(C+D) / (A+B)]
  - last 50% of trials 
- A, B, C, D
  - 100% of trials
  - each 10% of trials 
  - each 20% of trials 
  - each 25% of trials 
  - each 33% of trials 
  - 100% of trials dichotomised to 60+ selections from this deck.  
- Modal deck selection (100% of trials)
- A+C
  - 100% of trials 
  - every 25% 
- C-A
  - 100% of trials
- D-B  
  - 100% of trials 
- Number of deck switches 
  - 100% of trials
  - each 20% of trials 
  - each 25% of trials 
- Number of deck switches after a loss on A or B to C or D (100% of trials)
- Number of deck non-switches after a loss on A (100% of trials)
- Number of deck non-switches after a loss on C or D (100% of trials)
- Money won (100% of trials)
- Money lost (100% of trials)
- Money won - lost (100% of trials)
- One-lag autocorrelation between deck selections and overall value (wins-losses) of the preceding card chosen for block 3, within 3 blocks
  - first 33.33% of trials
  - last 33.33% of trials
  - difference between first and last third of trials

## Deck choice counts

### All trials

- [(C+D)-(A+B)]
  - dichotomised to less than 0 (impaired) or more than 0 (not impaired)
  - trichotomised to less than -10 (impaired), more than 10 (not impaired), or otherwise borderline
  - trichotomised to binomial test p-value for each participant using 100% of trials; significant disadvantageous choices, significant advantageous choices, and non-significant.
  - normalised via t-scores, 100% of trials [linear transformation of [(C+D)-(A+B)], not retained]
  - normalised via t-scores and dichotomised using t-score = 39, 100% of trials
  - normalised via t-scores and dichotomised using t-score = 50, 100% of trials
- C+D and logit transformed
- A, B, C, D
  - 100% of trials dichotomised to 60+ selections from this deck.
- A+C
- C-A
- D-B  
- Modal deck selection (100% of trials) [see note below]

Note that the method of using a binomial test can be simplified to whether the number of C+D choices is > `r qbinom(0.975, size = 100, prob = 0.5)` (scored as 1), < `r qbinom(0.025, size = 100, prob = 0.5)` (scored as -1), and otherwise scored as 0.

Note that one study in the literature calculated modal deck choice (Mapelli et al. 2004) and used this appropriately in a chi-square test. However, this categorical variable cannot be used in our multiverse. As compromise, we calculate the modal deck and then dichotomise this as 0 (disadvantageous, A or B) or 1 (advantageous, C or D).

```{r}

data_scored_deck_choices_temp <- data_processed_trial_level |>
  group_by(id) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total,
         C_plus_D = C+D,
         A_plus_B = A+B) |>
  select(-total) |>
  mutate(A_60_cutoff = as.numeric(A >= .60),
         B_60_cutoff = as.numeric(B >= .60),
         C_60_cutoff = as.numeric(C >= .60),
         D_60_cutoff = as.numeric(D >= .60),
         A_plus_C  = A+C, 
         C_minus_A = C-A, 
         C_plus_D_log_transform = log(C+D / (1 - C+D)),
         D_minus_B = D-B, 
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B,
         C_plus_D_minus_A_plus_B_binomial_significance = as.numeric(case_when(C_plus_D*100 > qbinom(0.975, size = 100, prob = 0.5) ~ 1,
                                                                              C_plus_D*100 < qbinom(0.025, size = 100, prob = 0.5) ~ -1,
                                                                              TRUE ~ 0)),
         C_plus_D_minus_A_plus_B_zero_cutoff = as.numeric(C_plus_D_minus_A_plus_B > 0),
         C_plus_D_minus_A_plus_B_absolute_10_cutoff = as.numeric(case_when(C_plus_D_minus_A_plus_B >= .10 ~ 1,
                                                                           C_plus_D_minus_A_plus_B <= .10 ~ -1,
                                                                           TRUE ~ 0)),
         C_plus_D_minus_A_plus_B_t_score = 50 + 10 * (C_plus_D_minus_A_plus_B - mean(C_plus_D_minus_A_plus_B)) / sd(C_plus_D_minus_A_plus_B),
         C_plus_D_minus_A_plus_B_t_score_39_cutoff = as.numeric(C_plus_D_minus_A_plus_B_t_score > 39),
         C_plus_D_minus_A_plus_B_t_score_50_cutoff = as.numeric(C_plus_D_minus_A_plus_B_t_score > 50)) |>
  # retain only some of these that were observed
  select(id,
         C_plus_D_minus_A_plus_B,
         C_plus_D_minus_A_plus_B_binomial_significance,
         C_plus_D_minus_A_plus_B_zero_cutoff,
         C_plus_D_minus_A_plus_B_absolute_10_cutoff,
         # the t score itself is not retained as it is a linear transformation of C_plus_D_minus_A_plus_B
         C_plus_D_minus_A_plus_B_t_score_39_cutoff,
         C_plus_D_minus_A_plus_B_t_score_50_cutoff,
         C_plus_D_log_transform, 
         A, B, C, D,
         A_60_cutoff, B_60_cutoff, C_60_cutoff, D_60_cutoff, 
         A_plus_C,
         C_minus_A, 
         D_minus_B)
         

get_mode_factor <- function(x) {
  tab <- table(x) # frequency table of deck choices
  levels(x)[which.max(tab)] # most frequent level, the first level wins ties
}

data_scored_deck_choices_modal <- data_processed_trial_level |>
  group_by(id) |>
  summarise(modal_deck_choice = get_mode_factor(choice), .groups = "drop") |>
  mutate(modal_deck_choice_dichotomised = case_when(modal_deck_choice %in% c("A", "B") ~ 0,
                                                    modal_deck_choice %in% c("C", "D") ~ 1)) |>
  select(-modal_deck_choice)

data_scored_deck_choices <- data_scored_deck_choices_temp |>
  left_join(data_scored_deck_choices_modal, by = "id") 

```

### By block 

- [(C+D)-(A+B)]
  - each 5% of trials 
  - each 17% of trials 
  - each 20% of trials 
  - each 20% of trials, normalised via t-scores [linear transformation of [(C+D)-(A+B)], not retained]
  - each 20% of trials, normalised via t-scores and dichotomised using t-score = 50 
  - each 20% of trials, normalised via t-scores and dichotomised using t-score = 39 
  - each 25% of trials 
  - each 33% of trials 
- A, B, C, D
  - each 10% of trials 
  - each 20% of trials 
  - each 25% of trials 
  - each 33% of trials 
- A+C
  - every 25% 
  
#### blocks of 5 trials

```{r}

data_scored_deck_choices_blocks_5_long <- data_processed_trial_level |>
  mutate(subset = ((trial - 1) %/% 5) + 1) |> # blocks of 5 trials
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(
    C_plus_D_minus_A_plus_B = (C+D)-(A+B),
  ) |>
  # retain only some based on what was observed in the meta methods review
  select(id, 
         subset, 
         C_plus_D_minus_A_plus_B)

data_scored_deck_choices_blocks_5 <- data_scored_deck_choices_blocks_5_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "C_plus_D_minus_A_plus_B_block5_")

```

#### blocks of 10 trials

```{r}

data_scored_deck_choices_blocks_10_long <- data_processed_trial_level |>
  mutate(subset = ((trial - 1) %/% 10) + 1) |> # blocks of 10 trials
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  # retain only some based on what was observed in the meta methods review
  select(id,
         subset,
         A, B, C, D)

data_scored_deck_choices_blocks_10 <- data_scored_deck_choices_blocks_10_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "block10_")

```

#### blocks of 16 trials

```{r}

data_scored_deck_choices_blocks_16_long <- data_processed_trial_level |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= .16 ~ 1,
                           trial_proportion > .16 & trial_proportion <= .33 ~ 2,
                           trial_proportion > .33 & trial_proportion <= .50 ~ 3,
                           trial_proportion > .50 & trial_proportion <= .67 ~ 4,
                           trial_proportion > .67 & trial_proportion <= .83 ~ 5,
                           trial_proportion > .83 ~ 6)) |>
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(
    C_plus_D_minus_A_plus_B = (C+D)-(A+B),
  ) |>
  # retain only some based on what was observed in the meta methods review
  select(id, 
         subset, 
         C_plus_D_minus_A_plus_B)

data_scored_deck_choices_blocks_16 <- data_scored_deck_choices_blocks_16_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "C_plus_D_minus_A_plus_B_block16_")

```

#### blocks of 20 trials

```{r}

data_scored_deck_choices_blocks_20_long <- data_processed_trial_level |>
  mutate(subset = ((trial - 1) %/% 20) + 1) |> # blocks of 20 trials
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total,
         C_plus_D_minus_A_plus_B = (C+D)-(A+B),
         C_plus_D_minus_A_plus_B_t_score = 50 + 10 * (C_plus_D_minus_A_plus_B - mean(C_plus_D_minus_A_plus_B)) / sd(C_plus_D_minus_A_plus_B),
         C_plus_D_minus_A_plus_B_t_score_39_cutoff = as.numeric(C_plus_D_minus_A_plus_B_t_score > 39),
         C_plus_D_minus_A_plus_B_t_score_50_cutoff = as.numeric(C_plus_D_minus_A_plus_B_t_score > 50)) |>
  select(-total) |>
  # retain only some based on what was observed in the meta methods review
  select(id, 
         subset, 
         C_plus_D_minus_A_plus_B,
         A, B, C, D, 
         C_plus_D_minus_A_plus_B_t_score,
         C_plus_D_minus_A_plus_B_t_score_39_cutoff,
         C_plus_D_minus_A_plus_B_t_score_50_cutoff)

data_scored_deck_choices_blocks_20 <- data_scored_deck_choices_blocks_20_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "block20_") |>
  # drop the block t scores, they are linear transformations of their C_plus_D_minus_A_plus_B equivalents
  select(-C_plus_D_minus_A_plus_B_t_score_block20_1,
         -C_plus_D_minus_A_plus_B_t_score_block20_2,
         -C_plus_D_minus_A_plus_B_t_score_block20_3,
         -C_plus_D_minus_A_plus_B_t_score_block20_4,
         -C_plus_D_minus_A_plus_B_t_score_block20_5)

```

#### blocks of 25 trials

```{r}

data_scored_deck_choices_blocks_25_long <- data_processed_trial_level |>
  mutate(subset = ((trial - 1) %/% 25) + 1) |> # blocks of 25 trials
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(
    C_plus_D_minus_A_plus_B = (C+D)-(A+B),
    A_plus_C = A+C
  ) |>
  # retain only some based on what was observed in the meta methods review
  select(id, 
         subset, 
         C_plus_D_minus_A_plus_B,
         A, B, C, D, 
         A_plus_C)

data_scored_deck_choices_blocks_25 <- data_scored_deck_choices_blocks_25_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "block25_")

```

#### blocks of 33 trials

```{r}

data_scored_deck_choices_blocks_33_long <- data_processed_trial_level |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= .33 ~ 1,
                           trial_proportion > .33 & trial_proportion <= .66 ~ 2,
                           trial_proportion > .66 ~ 3)) |>
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(
    C_plus_D_minus_A_plus_B = (C+D)-(A+B),
  ) |>
  # retain only some based on what was observed in the meta methods review
  select(id, 
         subset, 
         C_plus_D_minus_A_plus_B,
         A, B, C, D)

data_scored_deck_choices_blocks_33 <- data_scored_deck_choices_blocks_33_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "block33_")

```

#### Combine

```{r}

data_scored_deck_choices_blocks <- data_scored_deck_choices_blocks_5 |>
  full_join(data_scored_deck_choices_blocks_10, by = "id") |>
  full_join(data_scored_deck_choices_blocks_16, by = "id") |>
  full_join(data_scored_deck_choices_blocks_20, by = "id") |>
  full_join(data_scored_deck_choices_blocks_25, by = "id") |>
  full_join(data_scored_deck_choices_blocks_33, by = "id") |>
  # calculate diff scores
  mutate(C_plus_D_minus_A_plus_B_block16_6minus1 =
           C_plus_D_minus_A_plus_B_block16_6 -
           C_plus_D_minus_A_plus_B_block16_1,
         C_plus_D_minus_A_plus_B_block33_3minus1 =
           C_plus_D_minus_A_plus_B_block33_3 -
           C_plus_D_minus_A_plus_B_block33_1)

```

### Subsets

#### [(C+D)-(A+B)] subsets

- [(C+D)-(A+B)]
  - first 10% of trials 
  - first 40% of trials
  - first 50% of trials
  - first 60% of trials
  - 11-50% of trials 
  - 51-80% of trials 
  - last 20% of trials 
  - last 25% of trials 
  - last 40% of trials
  - last 50% of trials
  - last 60% of trials
  - diff between first 40% of trials and last 60% of trials
  - diff between first 33% of trials and last 33% of trials
  - diff between first 17% of trials and last 17% of trials 

```{r}

C_plus_D_minus_A_plus_B_subset <- function(dat, start, end){
  new_name <- paste0("C_plus_D_minus_A_plus_B_", start, "_to_", end)
  
  dat |>
    filter(trial %in% start:end) |>
    group_by(id) |>
    count(choice) |>
    pivot_wider(names_from = choice,
                values_from = n) |>
    ungroup() |>
    # replacing NA with 0
    mutate(A = ifelse(is.na(A), 0, A),
           B = ifelse(is.na(B), 0, B),
           C = ifelse(is.na(C), 0, C),
           D = ifelse(is.na(D), 0, D)) |>
    # convert counts to proportions to control for number of trials
    mutate(total = A+B+C+D,
           A = A/total,
           B = B/total,
           C = C/total,
           D = D/total) |>
    select(-total) |>
    mutate(
      C_plus_D_minus_A_plus_B = (C+D)-(A+B),
    ) |>
    select(-A, -B, -C, -D) |>
    rename(!!new_name := C_plus_D_minus_A_plus_B)
}

data_scored_deck_choices_subsets_1 <- 
  C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 10) |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 17), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 33), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 40), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 50), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 60), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 11, 50), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 41, 100), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 51, 80), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 51, 100), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 61, 100), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 68, 100), by = "id") |>
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 84, 100), by = "id") |>
  # calculate diff scores
  mutate(C_plus_D_minus_A_plus_B_last_60_minus_first_40 = 
           C_plus_D_minus_A_plus_B_41_to_100 - 
           C_plus_D_minus_A_plus_B_1_to_40,
         C_plus_D_minus_A_plus_B_last_33_minus_first_33 = 
           C_plus_D_minus_A_plus_B_68_to_100 - 
           C_plus_D_minus_A_plus_B_1_to_33,
         C_plus_D_minus_A_plus_B_last_17_minus_first_17 = 
           C_plus_D_minus_A_plus_B_84_to_100 - 
           C_plus_D_minus_A_plus_B_1_to_17) |>
  # drop subsets that duplicate block columns (1 to 33, 84 to 100) or only feed the diff scores (1 to 17)
  select(-C_plus_D_minus_A_plus_B_1_to_33,
         -C_plus_D_minus_A_plus_B_1_to_17,
         -C_plus_D_minus_A_plus_B_84_to_100)

```

#### [(C+D)-(A+B)] subsets dichotomised

- [(C+D)-(A+B)] dichotomised
  - dichotomised using median split, last 60% 
  - dichotomised using 0 cut-off, 1-60%
  - dichotomised using 0 cut-off, 61-100%

```{r}

data_scored_deck_choices_subsets_2 <- 
  # [(C+D)-(A+B)] dichotomised using median split, last 60% 
  C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 41, 100) |>
  mutate(C_plus_D_minus_A_plus_B_41_to_100_mediansplit = as.numeric(
    C_plus_D_minus_A_plus_B_41_to_100 > median(C_plus_D_minus_A_plus_B_41_to_100)
  )) |>
  select(-C_plus_D_minus_A_plus_B_41_to_100) |>
  
  # [(C+D)-(A+B)] dichotomised using 0 cut-off, 1-60%
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 1, 60), by = "id") |>
  mutate(C_plus_D_minus_A_plus_B_1_to_60_0cutuff = as.numeric(
    C_plus_D_minus_A_plus_B_1_to_60 > 0
  )) |>
  select(-C_plus_D_minus_A_plus_B_1_to_60) |>
  
  # [(C+D)-(A+B)] dichotomised using 0 cut-off, 61-100%
  full_join(C_plus_D_minus_A_plus_B_subset(data_processed_trial_level, 61, 100), by = "id") |>
  mutate(C_plus_D_minus_A_plus_B_61_to_100_0cutuff = as.numeric(
    C_plus_D_minus_A_plus_B_61_to_100 > 0
  )) |>
  select(-C_plus_D_minus_A_plus_B_61_to_100)

```

#### [(C+D)/(A+B)] subsets

- [(C+D)/(A+B)]
  - last 50% of trials 
  
Note that the original article that employed this strategy does not explicate how they handle Inf values, so I simply replace them with NA.

```{r}

data_scored_deck_choices_subsets_3 <- data_processed_trial_level |>
  filter(trial %in% 51:100) |> 
  group_by(id) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(
    C_plus_D_divided_A_plus_B = (C+D)/(A+B),
    C_plus_D_divided_A_plus_B = ifelse(is.infinite(C_plus_D_divided_A_plus_B), NA, C_plus_D_divided_A_plus_B)
  ) |>
  select(-A, -B, -C, -D)

```

#### Combine

```{r}

data_scored_deck_choices_subsets <- data_scored_deck_choices_subsets_1 |>
  full_join(data_scored_deck_choices_subsets_2, by = "id") |>
  full_join(data_scored_deck_choices_subsets_3, by = "id")

```

## Deck switching

### All trials

- Number of deck switches 
  - 100% of trials
- Number of deck switches after a loss on A or B to C or D (100% of trials)
- Number of deck non-switches after a loss on A (100% of trials)
- Number of deck non-switches after a loss on C or D (100% of trials)

```{r}

# "trials participants switch decks"
data_scored_deck_switching_any <- data_processed_trial_level |>
  group_by(id) |>
  mutate(switch = choice != dplyr::lag(choice)) |>
  summarize(proportion_switch_any = mean(switch, na.rm = TRUE))

# "trials participants switch from disadvantageous to advantageous decks after loss"
# proportion of trials after a loss with a switch between decks a or b and decks c or d, in either direction
data_scored_deck_switching_to_cd_after_loss_on_disadvantageous <- data_processed_trial_level |>
  mutate(choice_binary = case_when(choice %in% c("C", "D")  ~ "advantageous",
                                   choice %in% c("A", "B")  ~ "disadvantageous")) |>
  group_by(id) |>
  mutate(switch = choice_binary != dplyr::lag(choice_binary)) |>
  filter(dplyr::lag(loss_binary == 1) == TRUE) |>
  summarize(proportion_switch_advantageous_after_loss = mean(switch, na.rm = TRUE)) 

# proportion of all trials on which deck a is chosen right after a loss on deck a
data_scored_deck_choices_A_after_loss_on_A <- data_processed_trial_level |>
  group_by(id) |>
  mutate(A_after_loss_on_A = choice == "A" & dplyr::lag(choice) == "A" & dplyr::lag(loss) < 0) |>
  summarize(A_after_loss_on_A = mean(A_after_loss_on_A, na.rm = TRUE)) |>
  ungroup()

# proportion of all trials on which deck c or d is chosen right after a loss on deck c or d
data_scored_deck_choices_CD_after_loss_on_CD <- data_processed_trial_level |>
  group_by(id) |>
  mutate(C_or_D_after_loss_on_C_or_D = choice %in% c("C", "D") & dplyr::lag(choice) %in% c("C", "D") & dplyr::lag(loss) < 0) |>
  summarize(C_or_D_after_loss_on_C_or_D = mean(C_or_D_after_loss_on_C_or_D, na.rm = TRUE)) |>
  ungroup()

# combine
data_scored_deck_switching <- data_scored_deck_switching_any |>
  full_join(data_scored_deck_switching_to_cd_after_loss_on_disadvantageous, by = "id") |>
  left_join(data_scored_deck_choices_A_after_loss_on_A, by = "id") |>
  left_join(data_scored_deck_choices_CD_after_loss_on_CD, by = "id")

```

### By block

- Number of deck switches 
  - each 20% of trials 
  - each 25% of trials 
  
#### blocks of 20 trials

```{r}

data_scored_deck_switching_any_blocks_20_long <- data_processed_trial_level |>
  mutate(subset = ((trial - 1) %/% 20) + 1) |> # blocks of 20 trials
  drop_na(subset) |>
  group_by(id, subset) |>
  mutate(switch = choice != dplyr::lag(choice)) |>
  summarize(proportion_switch_any = mean(switch, na.rm = TRUE)) |>
  ungroup()

data_scored_deck_switching_any_blocks_20 <- data_scored_deck_switching_any_blocks_20_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "proportion_switch_any_")

```

#### blocks of 25 trials

```{r}

data_scored_deck_switching_any_blocks_25_long <- data_processed_trial_level |>
  mutate(subset = ((trial - 1) %/% 25) + 1) |> # blocks of 25 trials
  drop_na(subset) |>
  group_by(id, subset) |>
  mutate(switch = choice != dplyr::lag(choice)) |>
  summarize(proportion_switch_any = mean(switch, na.rm = TRUE)) |>
  ungroup()

data_scored_deck_switching_any_blocks_25 <- data_scored_deck_switching_any_blocks_25_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset),
              names_prefix = "proportion_switch_any__")

```

#### Combine

```{r}

data_scored_deck_switching_blocks <- data_scored_deck_switching_any_blocks_20 |>
  full_join(data_scored_deck_switching_any_blocks_25, by = "id") 

```

## Total money won/lost/balance

Ignores starting capital

- Money won (100% of trials)
- Money lost (100% of trials)
- Money won - lost (100% of trials)

```{r}

data_scored_earnings <- data_processed_trial_level |>
  group_by(id) |>
  summarize(total_won = sum(win),
            total_lost = sum(loss),
            total_balance = sum(balance_for_trial))

```

## One-lag autocorrelation between deck selections and overall value (wins-losses) of the preceding card

### By block (33% subsets) and difference between first and last block

- One-lag autocorrelation between deck selections and overall value (wins-losses) of the preceding card chosen for block 3, within 3 blocks
  - first 33.33% of trials
  - last 33.33% of trials
  - difference between first and last third of trials

```{r}

data_scored_one_lag_autocorrelation <- data_processed_trial_level %>%
  mutate(block = case_when(trial < 34 ~ 1,
                           trial > 67 ~ 3,
                           TRUE ~ 2)) |>
  group_by(id, block) %>%
  mutate(deck_choice_dichotomous = case_when(choice %in% c("A", "B") ~ 0,
                                             choice %in% c("C", "D") ~ 1),
         lagged_balance_for_trial = lag(balance_for_trial)) %>%
  summarise(one_lag_autocorrelation = cor(deck_choice_dichotomous, lagged_balance_for_trial, use = "complete.obs"), 
            .groups = "drop") |>
  pivot_wider(names_from = "block", 
              values_from = "one_lag_autocorrelation",
              names_prefix = "one_lag_autocorrelation_block_") |>
  select(-one_lag_autocorrelation_block_2) |>
  mutate(one_lag_autocorrelation_block_3_minus_block_1 = ifelse(!is.na(one_lag_autocorrelation_block_1) & !is.na(one_lag_autocorrelation_block_3),
                                                                one_lag_autocorrelation_block_3 - one_lag_autocorrelation_block_1,
                                                                NA))

```

## Combine

```{r}

data_processed_participant_level <- data_scored_deck_choices |>
  full_join(data_scored_deck_choices_blocks, by = "id") |>
  full_join(data_scored_deck_choices_subsets, by = "id") |>

  full_join(data_scored_deck_switching, by = "id") |>
  full_join(data_scored_deck_switching_blocks, by = "id") |>
  
  full_join(data_scored_earnings, by = "id") |>

  full_join(data_scored_one_lag_autocorrelation, by = "id") |>
  
  # round all
  round_df(4)

```

## Write to disk

```{r}

write_csv(data_processed_participant_level,
          "../../data/processed/multiverse/data_processed_participant_level.csv")

```

# Session Info

```{r session-info}

sessionInfo()

```
