---
title: "IGT multiverse"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# TODO

- correct wins/losses scoring methods for number of trials in the task


```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

# Get data

```{r}

# dependencies
library(tidyverse)
library(knitr)
library(kableExtra)
#library(sjPlot)
#library(forcats)
library(janitor)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, janitor::round_half_up, digits = n_digits)
}

# get data
load("../../data/raw/multiverse/IGTdata.rdata")

```

# Score items

## Win/loss

```{r}

data_wins <- 
  bind_rows(
    wi_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("100trials", id, sep = "_")),
    wi_95 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("95trials", id, sep = "_")),
    wi_150 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("150trials", id, sep = "_"))
  )

data_losses <- 
  bind_rows(
    lo_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("100trials", id, sep = "_")),
    lo_95 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("95trials", id, sep = "_")),
    lo_150 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("150trials", id, sep = "_"))
  )

data_winloss <- 
  full_join(data_wins, data_losses, by = c("id", "trial")) |>
  # dichotomise - TODO check logic 
  mutate(balance = win + loss, 
         win_binary = ifelse(win > 50, 1, 0),  # TODO is > 50 right here?
         loss_binary = ifelse(loss < 0, 1, 0),
         advantagous = ifelse(balance > 0, 1, 0),
         disadvantagous = ifelse(balance < 0, 1, 0))

# check that no advantagous==1 & disadvantagous==1 exist
data_winloss |>
  count(advantagous, disadvantagous)

```

## Choices

This assumes 1 = A, 2 = B etc. Check against the paper?

```{r}

data_choice <- 
  bind_rows(
    choice_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("100trials", id, sep = "_")),
    choice_95 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("95trials", id, sep = "_")),
    choice_150 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("150trials", id, sep = "_"))
  ) |>
  mutate(choice = as.factor(case_when(choice == 1 ~ "A",
                                      choice == 2 ~ "B",
                                      choice == 3 ~ "C",
                                      choice == 4 ~ "D")))

```

## Combine

```{r}

data_processed_trial_level <- data_winloss |>
  full_join(data_choice, by = c("id", "trial"))

```

## Compare decks

To check that I have correctly labelled the disadvantagous decks (A and B) and the advantagous ones (C and D)

```{r}

data_processed_trial_level |>
  group_by(choice) |>
  summarize(mean_win = mean(win),
            mean_loss = mean(loss),
            mean_balance = mean(balance)) |>
  round_df(0)

```

## Write to disk

```{r}

write_csv(data_processed_trial_level, "../../data/processed/multiverse/data_processed_trial_level.csv")

```

# Score participants

## (Dis)advantageous choices 

### All trials

```{r}

# TODO MAP THESE ONTO SCORING STRATEGIES, NOT SURE IF NEEDED

# TODO need to correct these for number of trials
# data_scored_advantage <- data_processed_trial_level |>
#   group_by(id) |>
#   summarize(mean_advantagous = mean(advantagous),
#             mean_disadvantagous = mean(disadvantagous)) |>
#   ungroup() |>
#   mutate(diff = mean_advantagous - mean_disadvantagous)
# 
# data_scored_advantage |>
#   select(-id) |>
#   cor() |>
#   round(2)
# 
# data_scored_advantage |>
#   pivot_longer(cols = -id) |>
#   ggplot(aes(value)) +
#   geom_density() +
#   facet_wrap(~name)

```

### Early and late trials 

Not observed in literature, but plausible. 

```{r}

# TODO

```

### Difference between early and late trials

Observed in literature, but simple early and late are not. 

```{r}

# TODO

```

## Wins/losses

Ignores starting capital

### All trials

```{r}

# TODO need to correct these for number of trials
data_scored_winloss <- data_processed_trial_level |>
  group_by(id) |>
  summarize(total_won = sum(win),
            total_lost = sum(loss)) |>
  ungroup() |>
  mutate(total_won_minus_lost = total_won + total_lost)

```

### Early and late trials 

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

### Difference between early and late trials

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

## Deck choice counts

### All trials

```{r}

data_scored_deck_choices <- data_processed_trial_level |>
  group_by(id) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(C_plus_A  = C+A,
         C_minus_A = C-A,
         D_plus_B  = D+B,
         D_minus_B = D-B,
         C_plus_D  = C+D,
         A_plus_B  = A+B,
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B,
         C_plus_D_divided_by_A_plus_B = C_plus_D/A_plus_B)

```

### Early and late trials

```{r}

data_scored_deck_choices_subset_long <- data_processed_trial_level |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= 0.4 ~ "early",   # TODO cutoffs could be varied
                            trial_proportion >= 0.4 ~ "late")) |>
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(C_plus_A  = C+A,
         C_minus_A = C-A,
         D_plus_B  = D+B,
         D_minus_B = D-B,
         C_plus_D  = C+D,
         A_plus_B  = A+B,
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B) 

data_scored_deck_choices_subset <- data_scored_deck_choices_subset_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset))

```

### Difference between early and late trials

Inherits the proportion values used to create the early and late subsets from the early/late chunk above. 

```{r}

data_scored_deck_choices_subset_diff <- data_scored_deck_choices_subset_long |>
  pivot_longer(names_to = "variable",
               values_to = "score",
               cols = c(-id, -subset)) |>
  pivot_wider(names_from = subset, 
              values_from = score) |>
  mutate(diff = late - early) |>
  select(-early, -late) |>
  pivot_wider(names_from = variable,
              values_from = diff,
              names_prefix = "diff_")

```

## Deck switching

### All trials

```{r}

# "trials participants switch decks"
data_scored_deck_switching_any <- data_processed_trial_level |>
  group_by(id) |>
  mutate(switch_any = choice == lag(choice)) |>
  summarize(proportion_switch_any = mean(switch_any, na.rm = TRUE))

# "deck switching after loss"
# proportion of switching vs non switching after loss, excluding trial on which there were not losses. 
# TODO This assumption of excluding non-loss trials needs checking.
data_scored_deck_switching_after_loss <- data_processed_trial_level |>
  filter(loss_binary == TRUE) |>
  group_by(id) |>
  mutate(switch_after_loss = choice == lag(choice)) |>
  summarize(proportion_switch_after_loss = mean(switch_after_loss, na.rm = TRUE))  

# "card selections from deck A which immediately followed a loss on that deck"
# proportion of switching away from deck A after a loss on deck A
data_scored_deck_switching_a_after_loss_on_a <- data_processed_trial_level |>
  filter(lag(loss_binary) == TRUE & lag(choice == "A")) |>
  group_by(id) |>
  mutate(switch_from_a_after_loss_on_a = choice != "A") |>
  summarize(proportion_switch_from_a_after_loss_on_a = mean(switch_from_a_after_loss_on_a, na.rm = TRUE))

# "trials participants switch from disadvantageous to advantageous decks after loss (total)"
# proportion of switching to deck C or D after a loss on A or B
data_scored_deck_switching_to_cd_after_loss_on_ab <- data_processed_trial_level |>
  filter(lag(loss_binary) == TRUE & lag(choice %in% c("A", "B"))) |>
  group_by(id) |>
  mutate(switch_to_cd_after_loss_on_ab = choice %in% c("C", "D")) |>
  summarize(proportion_switch_to_cd_after_loss_on_ab = mean(switch_to_cd_after_loss_on_ab, na.rm = TRUE))
       
# combine
data_scored_deck_switching <- data_scored_deck_switching_any |>
  full_join(data_scored_deck_switching_after_loss, by = "id") |>
  full_join(data_scored_deck_switching_a_after_loss_on_a, by = "id") |>
  full_join(data_scored_deck_switching_to_cd_after_loss_on_ab, by = "id")

```

### Early and late trials

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

### Difference between early and late trials

Not observed in literature, but plausible. To implement or not?

```{r}

# TODO

```

## Combine

```{r}

data_processed_participant_level <- data_scored_winloss |>
  full_join(data_scored_deck_choices,             by = "id") |>
  full_join(data_scored_deck_choices_subset,      by = "id") |>
  full_join(data_scored_deck_choices_subset_diff, by = "id") |>
  full_join(data_scored_deck_switching,           by = "id")

```

## Write to disk

```{r}

write_csv(data_processed_participant_level, "../../data/processed/multiverse/data_processed_participant_level.csv")

```



