---
title: "IGT multiverse"
subtitle: "Data processing"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

# Get data

```{r}

options(scipen=999)

# dependencies
library(tidyverse)
library(knitr)
library(kableExtra)
#library(sjPlot)
#library(forcats)
library(janitor)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, janitor::round_half_up, digits = n_digits)
}

# get data
load("../../data/raw/multiverse/IGTdata.rdata")

```

# Score items

## Win/loss

```{r}

data_wins <- 
  bind_rows(
    wi_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "win") |>
      mutate(trial = as.numeric(str_remove(trial, "Wins_")),
             id = paste("100trials", id, sep = "_")),
    # # i originally included the data for the other procedural versions of the task, but this greatly complicates things as you have to standardize things like money won by the number of trials. better to simply not include these.
    # wi_95 |>
    #   as.tibble() |>
    #   rownames_to_column(var = "id") |>
    #   pivot_longer(cols = -id,
    #                names_to = "trial",
    #                values_to = "win") |>
    #   mutate(trial = as.numeric(str_remove(trial, "Wins_")),
    #          id = paste("95trials", id, sep = "_")),
    # wi_150 |>
    #   as.tibble() |>
    #   rownames_to_column(var = "id") |>
    #   pivot_longer(cols = -id,
    #                names_to = "trial",
    #                values_to = "win") |>
    #   mutate(trial = as.numeric(str_remove(trial, "Wins_")),
    #          id = paste("150trials", id, sep = "_"))
  )

data_losses <- 
  bind_rows(
    lo_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "loss") |>
      mutate(trial = as.numeric(str_remove(trial, "Losses_")),
             id = paste("100trials", id, sep = "_")),
    # # i originally included the data for the other procedural versions of the task, but this greatly complicates things as you have to standardize things like money won by the number of trials. better to simply not include these.
    # lo_95 |>
    #   as.tibble() |>
    #   rownames_to_column(var = "id") |>
    #   pivot_longer(cols = -id,
    #                names_to = "trial",
    #                values_to = "loss") |>
    #   mutate(trial = as.numeric(str_remove(trial, "Losses_")),
    #          id = paste("95trials", id, sep = "_")),
    # lo_150 |>
    #   as.tibble() |>
    #   rownames_to_column(var = "id") |>
    #   pivot_longer(cols = -id,
    #                names_to = "trial",
    #                values_to = "loss") |>
    #   mutate(trial = as.numeric(str_remove(trial, "Losses_")),
    #          id = paste("150trials", id, sep = "_"))
  )

data_winloss <- 
  full_join(data_wins, data_losses, by = c("id", "trial")) |>
  mutate(balance_for_trial = win + loss, 
         loss_binary = ifelse(loss < 0, 1, 0))

```

## Choices

This assumes 1 = A, 2 = B, 3 = C, 4 = D. This seems to be correct when comparing against the average amount won per deck.

```{r}

data_choice <- 
  bind_rows(
    choice_100 |>
      as.tibble() |>
      rownames_to_column(var = "id") |>
      pivot_longer(cols = -id,
                   names_to = "trial",
                   values_to = "choice") |>
      mutate(trial = as.numeric(str_remove(trial, "Choice_")),
             id = paste("100trials", id, sep = "_")),
    # choice_95 |>
    #   as.tibble() |>
    #   rownames_to_column(var = "id") |>
    #   pivot_longer(cols = -id,
    #                names_to = "trial",
    #                values_to = "choice") |>
    #   mutate(trial = as.numeric(str_remove(trial, "Choice_")),
    #          id = paste("95trials", id, sep = "_")),
    # choice_150 |>
    #   as.tibble() |>
    #   rownames_to_column(var = "id") |>
    #   pivot_longer(cols = -id,
    #                names_to = "trial",
    #                values_to = "choice") |>
    #   mutate(trial = as.numeric(str_remove(trial, "Choice_")),
    #          id = paste("150trials", id, sep = "_"))
  ) |>
  mutate(choice = as.factor(case_when(choice == 1 ~ "A",
                                      choice == 2 ~ "B",
                                      choice == 3 ~ "C",
                                      choice == 4 ~ "D")))

```

## Combine

```{r}

data_processed_trial_level <- data_winloss |>
  full_join(data_choice, by = c("id", "trial"))

```

## Compare decks

To check that I have correctly labelled the disadvantagous decks (A and B) and the advantagous ones (C and D)

```{r}

data_processed_trial_level |>
  group_by(choice) |>
  summarize(mean_win = mean(win),
            mean_loss = mean(loss),
            mean_balance = mean(balance_for_trial)) |>
  round_df(0)

```

## Check all participants have 100 trials

```{r}

data_processed_trial_level |>
  count(id) |>
  count(n)

```

## Write to disk

```{r}

write_csv(data_processed_trial_level,
          "../../data/processed/multiverse/data_processed_trial_level.csv")

```

# Score participants

## Overview

Data is scored using the following factorial design:

- 3 X subset: 
  - all trials
  - first vs last block [20%]
  - difference between first and last block [20%]
- 21 X scoring approach: 
  - total money won, total lost, total final balance (won-lost)
  - number of trials on which a loss was incurred [brutto loss, agnostic to money won on that trial]
  - number of selection from deck(s) A, B, C, D, A_plus_B, A_plus_C, B_plus_D, C_minus_A, C_plus_D, C_minus_D, D_minus_B, C_plus_D_minus_A_plus_B, C_plus_D_divided_by_A_plus_B.
  - number of switches between decks: total number of switches between decks, total number of switches between decks on trials following a loss trial, total number of switches between advantageous vs. disadvantageous decks (C and D vs A and B), total number of switches between advantageous vs. disadvantageous decks (C and D vs A and B) on trials following a loss trial.
- total of 63 scores per person.

Note that scoring strategies based on reactiont times are also sometimes used, but the dataset we were able to obtain did not contain reaction time data so this was not possible.

Other scoring strategies that could be implemented which were observed in the literature:

- [(C+D)-(A+B)]: each 5% of trials, first 40% of trials, first 50% of trials, first 60% of trials, last 40% of trials, last 50% of trials, last 60% of trials, diff between first 40% of trials and last 60% of trials, diff between first 33% of trials and last 33% of trials, dichotomised to less than -10 (impaired) or more than -10 (not impaired), dichotomised to less than 0 (impaired) or more than 0 (not impaired), last 60% dichotomised using median split.  [done in spreadsheet]
- A, B, C, D: first 33%, second 33%, third 33%, dichotomised to 60+ selections from this deck.  [done in spreadsheet]
- A+B: first 10%, trials 11-50, trials 51-80, first 33%, second 33%, third 33%  [done in spreadsheet]
- C+D: first 10%, trials 11-50, trials 51-80, each 16.66%, first 16.66% of trials vs last 16.66% of trials [done in spreadsheet]
- A, B, C, D, C+D, B+D [weird]: first 25%, second 25%, third 25%, fourth 25%  [done in spreadsheet]
- C-A: 100% of trials [done in spreadsheet]
- D-B: 100% of trials [done in spreadsheet]
- A+C: 100% [done in spreadsheet] *NOT IMPLEMENTED BELOW?*
- B+D: [weird]: 100%, first 25%, second 25%, third 25%, fourth 25% [done in spreadsheet] *NOT IMPLEMENTED BELOW?*

- not switching decks to disadvantageous decks after receiving a punishment on an advantageous deck
- A following a loss on A

Other rare strategies

- one lag autocorrelation between deck selections and overall value (wins-losses) of the preceding card chosen for block 3, within 3 blocks (first 33.33$, last 33.33% of trials)

## Deck choice counts

### All trials

```{r}

data_scored_deck_choices <- data_processed_trial_level |>
  group_by(id) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(A_plus_B  = A+B, 
         A_plus_C  = A+C, 
         B_plus_D  = B+D, 
         C_minus_A = C-A, 
         C_plus_D  = C+D, 
         C_minus_D = C-D, 
         D_minus_B = D-B, 
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B,
         C_plus_D_divided_by_A_plus_B = C_plus_D/A_plus_B) # TODO does this actually appear in the literature??

```

### By block (20% subsets)

```{r}

data_scored_deck_choices_subset_long <- data_processed_trial_level |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= .20 ~ 1,
                           trial_proportion > .20 & trial_proportion <= .40 ~ 2,
                           trial_proportion > .40 & trial_proportion <= .60 ~ 3,
                           trial_proportion > .60 & trial_proportion <= .80 ~ 4,
                           trial_proportion > .80 ~ 5)) |>
  drop_na(subset) |>
  group_by(id, subset) |>
  count(choice) |>
  pivot_wider(names_from = choice,
              values_from = n) |>
  ungroup() |>
  # replacing NA with 0
  mutate(A = ifelse(is.na(A), 0, A),
         B = ifelse(is.na(B), 0, B),
         C = ifelse(is.na(C), 0, C),
         D = ifelse(is.na(D), 0, D)) |>
  # convert counts to proportions to control for number of trials
  mutate(total = A+B+C+D,
         A = A/total,
         B = B/total,
         C = C/total,
         D = D/total) |>
  select(-total) |>
  mutate(A_plus_B  = A+B, 
         A_plus_C  = A+C, 
         B_plus_D  = B+D, 
         C_minus_A = C-A, 
         C_plus_D  = C+D, 
         C_minus_D = C-D,
         D_minus_B = D-B, 
         C_plus_D_minus_A_plus_B = C_plus_D-A_plus_B,
         C_plus_D_divided_by_A_plus_B = C_plus_D/A_plus_B)

data_scored_deck_choices_subset <- data_scored_deck_choices_subset_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset))

```

### Difference between first and last block

Inherits the proportion values used to create the early and late subsets from the early/late chunk above. 

```{r}

data_scored_deck_choices_subset_diff <- data_scored_deck_choices_subset_long |>
  pivot_longer(names_to = "variable",
               values_to = "score",
               cols = c(-id, -subset)) |>
  pivot_wider(names_from = subset, 
              values_from = score,
              names_prefix = "block_") |>
  mutate(diff = block_5 - block_1) |>
  select(id, variable, diff) |>
  pivot_wider(names_from = variable, 
              values_from = diff) |>
  rename(A_block_5_minus_block_1 = A, 
         B_block_5_minus_block_1 = B, 
         C_block_5_minus_block_1 = C, 
         D_block_5_minus_block_1 = D, 
         A_plus_B_block_5_minus_block_1 = A_plus_B, 
         A_plus_C_block_5_minus_block_1 = A_plus_C, 
         B_plus_D_block_5_minus_block_1 = B_plus_D, 
         C_minus_A_block_5_minus_block_1 = C_minus_A, 
         C_plus_D_block_5_minus_block_1 = C_plus_D, 
         C_minus_D_block_5_minus_block_1 = C_minus_D, 
         D_minus_B_block_5_minus_block_1 = D_minus_B, 
         C_plus_D_minus_A_plus_B_block_5_minus_block_1 = C_plus_D_minus_A_plus_B, 
         C_plus_D_divided_by_A_plus_B_block_5_minus_block_1 = C_plus_D_divided_by_A_plus_B)

```

## Total money won/lost/balance

Ignores starting capital

### All trials

```{r}

data_scored_earnings <- data_processed_trial_level |>
  group_by(id) |>
  summarize(total_won = sum(win),
            total_lost = sum(loss),
            total_balance = sum(balance_for_trial))

```

### By block (20% subsets)

```{r}

data_scored_earnings_subset_long <- data_processed_trial_level |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= .20 ~ 1,
                            trial_proportion > .20 & trial_proportion <= .40 ~ 2,
                            trial_proportion > .40 & trial_proportion <= .60 ~ 3,
                            trial_proportion > .60 & trial_proportion <= .80 ~ 4,
                            trial_proportion > .80 ~ 5)) |>
  group_by(id, subset) |>
  summarize(total_won = sum(win),
            total_lost = sum(loss),
            total_balance = sum(balance_for_trial))

data_scored_earnings_subset <- data_scored_earnings_subset_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset))

```

### Difference between first and last block

```{r}

data_scored_earnings_subset_diff <- data_scored_earnings_subset_long |>
  pivot_longer(names_to = "variable",
               values_to = "score",
               cols = c(-id, -subset)) |>
  pivot_wider(names_from = subset, 
              values_from = score,
              names_prefix = "block_") |>
  mutate(earnings_diff = block_5 - block_1) |>
  select(id, variable, earnings_diff) |>
  pivot_wider(names_from = variable,
              values_from = earnings_diff) |>
  rename(total_won_block_5_minus_block_1 = total_won,
         total_lost_block_5_minus_block_1 = total_lost,
         total_balance_block_5_minus_block_1 = total_balance)

```

## Loss trials

### All trials

```{r}

data_scored_loss_trials <- data_processed_trial_level |>
  group_by(id) |>
  summarize(total_loss_trials = sum(loss_binary))

```

### By block (20% subsets)

```{r}

data_scored_loss_trials_subset_long <- data_processed_trial_level |>
  group_by(id) |>
  mutate(total_trials = max(trial),
         trial_proportion = trial/total_trials) |>
  ungroup() |>
  mutate(subset = case_when(trial_proportion <= .20 ~ 1,
                            trial_proportion > .20 & trial_proportion <= .40 ~ 2,
                            trial_proportion > .40 & trial_proportion <= .60 ~ 3,
                            trial_proportion > .60 & trial_proportion <= .80 ~ 4,
                            trial_proportion > .80 ~ 5)) |>
  group_by(id, subset) |>
  summarize(total_loss_trials = sum(loss_binary))

data_scored_loss_trials_subset <- data_scored_loss_trials_subset_long |>
  pivot_wider(names_from = "subset",
              values_from = c(-id, -subset))

```

### Difference between first and last block

```{r}

data_scored_loss_trials_subset_diff <- data_scored_loss_trials_subset_long |>
  pivot_wider(names_from = subset, 
              values_from = total_loss_trials,
              names_prefix = "block_") |>
  mutate(total_loss_trials_block_5_minus_block_1 = block_5 - block_1) |>
  select(id, total_loss_trials_block_5_minus_block_1)

```


## Deck switching

### All trials

```{r}

# "trials participants switch decks"
data_scored_deck_switching_any <- data_processed_trial_level |>
  group_by(id) |>
  mutate(switch = choice != dplyr::lag(choice)) |>
  summarize(proportion_switch_any = mean(switch, na.rm = TRUE))

# "deck switching after loss"
# proportion of switching vs non switching after loss, excluding trial on which there were not losses. 
data_scored_deck_switching_after_loss <- data_processed_trial_level |>
  group_by(id) |>
  mutate(switch = choice != dplyr::lag(choice)) |>
  filter(dplyr::lag(loss_binary == 1) == TRUE) |>
  summarize(proportion_switch_after_loss = mean(switch, na.rm = TRUE)) 

# "trials participants switch from disadvantageous to advantageous decks"
data_scored_deck_switching_to_cd_any <- data_processed_trial_level |>
  mutate(choice_binary = case_when(choice %in% c("C", "D")  ~ "advantageous",
                                   choice %in% c("A", "B")  ~ "disadvantageous")) |>
  group_by(id) |>
  mutate(switch = choice_binary != dplyr::lag(choice_binary)) |>
  summarize(proportion_switch_advantageous = mean(switch, na.rm = TRUE))

# "trials participants switch from disadvantageous to advantageous decks after loss"
# proportion of switching to deck C or D after a loss on A or B
data_scored_deck_switching_to_cd_after_loss_on_disadvantageous <- data_processed_trial_level |>
  mutate(choice_binary = case_when(choice %in% c("C", "D")  ~ "advantageous",
                                   choice %in% c("A", "B")  ~ "disadvantageous")) |>
  group_by(id) |>
  mutate(switch = choice_binary != dplyr::lag(choice_binary)) |>
  filter(dplyr::lag(loss_binary == 1) == TRUE) |>
  summarize(proportion_switch_advantageous_after_loss = mean(switch, na.rm = TRUE)) 
       
# combine
data_scored_deck_switching <- data_scored_deck_switching_any |>
  full_join(data_scored_deck_switching_after_loss, by = "id") |>
  full_join(data_scored_deck_switching_to_cd_any, by = "id") |>
  full_join(data_scored_deck_switching_to_cd_after_loss_on_disadvantageous, by = "id")

```

### By block (20% subsets)

Add?

### Difference between first and last block

Add?

## Combine

```{r}

data_combined <- data_scored_earnings |>
  full_join(data_scored_earnings_subset, by = "id") |>
  full_join(data_scored_earnings_subset_diff, by = "id") |>
  
  full_join(data_scored_loss_trials, by = "id") |>
  full_join(data_scored_loss_trials_subset, by = "id") |>
  full_join(data_scored_loss_trials_subset_diff, by = "id") |>
  
  full_join(data_scored_deck_choices, by = "id") |>
  full_join(data_scored_deck_choices_subset, by = "id") |>
  full_join(data_scored_deck_choices_subset_diff, by = "id") |>

  full_join(data_scored_deck_switching, by = "id")
  # full_join(data_scored_deck_switching_subset, by = "id") |> # add?
  # full_join(data_scored_deck_switching_subset_diff, by = "id") # add?

```

#### Check my intutiion that some of these are monotonic / redundant with one another

```{r}

monotonic_scoring_methods <- data_combined |>
  select(-id) |>
  select(!ends_with("_1") & !ends_with("_2") & !ends_with("_3") & !ends_with("_4") & !ends_with("_5")) |>
  cor() |>
  as.data.frame() |>
  rownames_to_column(var = "var1") |>
  pivot_longer(cols = -var1,
               names_to = "var2",
               values_to = "r") |>
  filter(abs(r) == 1 & 
           var1 != var2)

monotonic_scoring_methods


data_combined |>
  select(A_plus_B, C_plus_D, C_plus_D_minus_A_plus_B, C_plus_D_divided_by_A_plus_B) |>
  cor()

data_combined |>
  select(B_plus_D, A_plus_C) |>
  cor()

```

- A_plus_B, C_plus_D, and C_plus_D_minus_A_plus_B are monotonic / redundant
- B_plus_D and A_plus_C are monotonic / redundant

# Averaged metrics

```{r}

data_processed_participant_level <- data_combined |>
  rowwise() |>
  mutate(
    # deck selections [(C+D)-(A+B)] for blocks 3, 4, and 5 (5 blocks)
    C_plus_D_minus_A_plus_B_345 = (C_plus_D_minus_A_plus_B_3 + C_plus_D_minus_A_plus_B_4 + C_plus_D_minus_A_plus_B_5)/3,
    # deck selections [(C+D)-(A+B)] for blocks 4 and 5 (5 blocks)
    C_plus_D_minus_A_plus_B_45 = (C_plus_D_minus_A_plus_B_4 + C_plus_D_minus_A_plus_B_5)/2,
    # deck selections [(C+D)-(A+B)] for blocks 1 and 2 (5 blocks)
    C_plus_D_minus_A_plus_B_12 = (C_plus_D_minus_A_plus_B_1 + C_plus_D_minus_A_plus_B_2)/2,
  ) |>
  ungroup()

```

## Write to disk

```{r}

write_csv(data_processed_participant_level, "../../data/processed/multiverse/data_processed_participant_level.csv")

data_processed_participant_level_frequent <- data_processed_participant_level |>
  select(id,
         C_plus_D_minus_A_plus_B,
         C_plus_D_minus_A_plus_B_1, 
         C_plus_D_minus_A_plus_B_2, 
         C_plus_D_minus_A_plus_B_3, 
         C_plus_D_minus_A_plus_B_4, 
         C_plus_D_minus_A_plus_B_5,
         # C_plus_D, # correlates 1.0 with C_plus_D_minus_A_plus_B
         A,
         B, 
         D, 
         C, 
         # A_plus_B, # correlates 1.0 with C_plus_D_minus_A_plus_B
         C_plus_D_minus_A_plus_B_345,
         #A_plus_B_5, 
         #C_plus_D_5, 
         # Attention to losses and gains parameter (EV model)
         # Recency parameter (EV model)
         # Choice consistency parameter (EV model parameter)
         C_plus_D_minus_A_plus_B_12,
         #A_plus_B_1, 
         #A_plus_B_2, 
         #A_plus_B_3, 
         #A_plus_B_4, 
         #C_plus_D_1, 
         #C_plus_D_2, 
         #C_plus_D_3, 
         #C_plus_D_4, 
         A_plus_C, 
         #B_plus_D,  # correlates 1.0 with A_plus_C
         C_plus_D_minus_A_plus_B_45,
         total_won_minus_lost,
         C_plus_D_minus_A_plus_B_block_5_minus_block_1) # this one isn't in the most common scores, but represents how two common scores are used in the analyses

write_csv(data_processed_participant_level_frequent, "../../data/processed/multiverse/data_processed_participant_level_frequent.csv")

common_cors <- 
  data_processed_participant_level_frequent |>
  select(-id) |>
  cor() |>
  round(2)

diag(common_cors) <- NA

common_cors[lower.tri(common_cors, diag = TRUE)] <- NA
abs(common_cors)

```


```{r}

top_loading_cors <- 
  data_processed_participant_level_frequent |>
  select(C_plus_D_minus_A_plus_B, # long term winning strategy
         total_won_minus_lost, # long term strategy of another sort
         C_plus_D_minus_A_plus_B_block_5_minus_block_1, # learning across task. returns the same score for people who start and continue well as those who start and continue badly.
         A_plus_C, # preference for frequent rewards over other factors
         A) |> # preference for frequent rewards despite long term large losses
  cor() |>
  round(2)

diag(top_loading_cors) <- NA

top_loading_cors[lower.tri(top_loading_cors, diag = TRUE)] <- NA
abs(top_loading_cors)

```



# scrap

```{r eval=FALSE, include=FALSE}

# # "card selections from deck A which immediately followed a loss on that deck"
# # proportion of switching away from deck A after a loss on deck A
# data_scored_deck_switching_a_after_loss_on_a <- data_processed_trial_level |>
#   filter(lag(loss_binary) == TRUE & lag(choice == "A")) |>
#   group_by(id) |>
#   mutate(switch_from_a_after_loss_on_a = choice != "A") |>
#   summarize(proportion_switch_from_a_after_loss_on_a = mean(switch_from_a_after_loss_on_a, na.rm = TRUE))

```
