---
title: "IGT scoring multiverse"
subtitle: "Analysis — Hayes et al. (2020) example dataset"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Packages

```{r packages}

library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)
library(patchwork)
library(corrr)

```

# Get data

```{r data}

data_processed_participant_level <- read_csv(
  "../../data/processed/example hayes2020/data_processed_participant_level.csv",
  show_col_types = FALSE
) %>%
  select(-id)

# identify and drop columns with no variance — correlate() fails on constants
zero_variance_cols <- data_processed_participant_level %>%
  summarise(across(everything(), ~ sd(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "col", values_to = "sd") %>%
  filter(sd == 0 | is.na(sd)) %>%
  pull(col)

data_for_cors <- data_processed_participant_level %>%
  select(-all_of(zero_variance_cols))

```

# Summary stats

Number of unique scoring methods implemented = `r ncol(data_processed_participant_level)`.

Number of scoring methods with sufficient variance for correlation = `r ncol(data_for_cors)`.

Zero-variance columns excluded from correlations: `r paste(zero_variance_cols, collapse = ", ")`.

Number of participants = `r nrow(data_processed_participant_level)`.

# Correlate scores

```{r correlate}

cors <- data_for_cors %>%
  correlate() %>%
  stretch() %>%
  filter(y != x) %>%  # remove diagonal
  arrange(desc(r))

```

## Perfect correlations (r = ±1)

In the original multiverse (Steingroever data), no pairs reached r = ±1 after removing redundant linear transforms during processing. This dataset may differ because of structural features of the Hayes 2020 task (e.g., fixed win amounts per deck: A/B always win 100, C/D always win 50 — which makes `total_won` a perfect linear function of choice proportion).

```{r perfect-cors}

# identify any pairs with |r| effectively equal to 1
perfect_cors <- cors %>%
  filter(x < y, abs(r) > 0.9999) %>%
  arrange(r)

if (nrow(perfect_cors) > 0) {
  perfect_cors %>%
    kable(digits = 4, caption = "Pairs with |r| = 1 (perfect linear dependence)") %>%
    kable_classic(full_width = FALSE)
} else {
  cat("No perfect correlations found.")
}

```

## Between all scoring methods

Note that pairs reaching r = ±1 represent perfect linear transformations of one another; see the table above for any such cases specific to this dataset.

```{r summary-all}

cors %>%
  filter(x < y) %>%  # upper triangle only
  summarize(
    n_correlations = n(),
    min_r          = min(r, na.rm = TRUE),
    max_r          = max(r, na.rm = TRUE),
    median_r       = median(r, na.rm = TRUE),
    mad_r          = mad(r, na.rm = TRUE),
    min_abs_r      = min(abs(r), na.rm = TRUE),
    max_abs_r      = max(abs(r), na.rm = TRUE),
    median_abs_r   = median(abs(r), na.rm = TRUE),
    mad_abs_r      = mad(abs(r), na.rm = TRUE)
  ) %>%
  mutate_if(is.numeric, round_half_up, digits = 3) %>%
  kable() %>%
  kable_classic(full_width = FALSE)

```

```{r plot-all}

# native correlations
p_native_all <- ggplot(cors %>% filter(x < y), aes(r)) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#0072B2") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  ggtitle("Correlations among all scoring strategies") +
  xlab("Correlation") +
  coord_cartesian(xlim = c(-1, 1)) +
  theme_bw()

# absolute correlations
p_absolute_all <- ggplot(cors %>% filter(x < y), aes(abs(r))) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#0072B2") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  xlab("Absolute correlation") +
  coord_cartesian(xlim = c(0, 1)) +
  theme_bw()

p_native_all
p_absolute_all

```

## Between modal scoring strategy and all others

The modal scoring strategy in the IGT literature is (C+D)−(A+B) across all trials.

```{r summary-modal}

cors %>%
  filter(x == "C_plus_D_minus_A_plus_B") %>%
  summarize(
    n_correlations = n(),
    min_r          = min(r, na.rm = TRUE),
    max_r          = max(r, na.rm = TRUE),
    median_r       = median(r, na.rm = TRUE),
    mad_r          = mad(r, na.rm = TRUE),
    min_abs_r      = min(abs(r), na.rm = TRUE),
    max_abs_r      = max(abs(r), na.rm = TRUE),
    median_abs_r   = median(abs(r), na.rm = TRUE),
    mad_abs_r      = mad(abs(r), na.rm = TRUE)
  ) %>%
  mutate_if(is.numeric, round_half_up, digits = 3) %>%
  kable() %>%
  kable_classic(full_width = FALSE)

```

```{r plot-modal}

p_native_modal <- ggplot(cors %>% filter(x == "C_plus_D_minus_A_plus_B"), aes(r)) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#E69F00") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  ggtitle("Correlations with modal scoring strategy") +
  xlab("Correlation") +
  coord_cartesian(xlim = c(-1, 1)) +
  theme_bw()

p_absolute_modal <- ggplot(cors %>% filter(x == "C_plus_D_minus_A_plus_B"), aes(abs(r))) +
  geom_histogram(binwidth = .1, boundary = 0, fill = "#E69F00") +
  scale_x_continuous(breaks = breaks_pretty(n = 10)) +
  scale_y_continuous(breaks = breaks_pretty(n = 7), name = "Frequency") +
  xlab("Absolute correlation") +
  coord_cartesian(xlim = c(0, 1)) +
  theme_bw()

p_native_modal
p_absolute_modal

```

## Combined plot

```{r plot-combined, fig.height=6, fig.width=9}

p_native_all +
  p_native_modal +
  p_absolute_all +
  p_absolute_modal +
  plot_layout(ncol = 2, nrow = 2)

ggsave("plot.pdf", width = 9, height = 6)
ggsave("plot.png", width = 9, height = 6, dpi = 600)

```

# Session info

```{r session-info}

sessionInfo()

```
