---
title: "Inter Rater Reliability Scoring"
author: "Annika Kuelpmann"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_download: true
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup}
#| include: false

# load knitr 
library(knitr)

# set knitr chunk options 
opts_chunk$set(echo = TRUE,
               message = FALSE,
               warning = FALSE)
```

# Load Packages

```{r packages}
#| echo: false

# load packages
library(readxl)
library(conflicted)
library(tidyverse)
library(irr)
library(boot)

# set preferred options 
conflicts_prefer(dplyr::filter)

```

# Data Preparation

```{r data prep}
#| echo: false

# read data files
data_r1 <- read_xlsx("../../data/raw/interrater reliability/codebook - rater 1/r1_scoring_aligned.xlsx")
data_r2 <- read_xlsx("../../data/raw/interrater reliability/codebook - rater 2/r2_scoring_aligned.xlsx")

# align data types
data_r1 <- data_r1 %>%
  mutate(score_count = as.numeric(score_count)) %>%
  mutate(score_id_new = as.character(score_id_new)) 
data_r2 <- data_r2 %>%
  mutate(score_count = as.numeric(score_count)) %>%
  mutate(score_id_new = as.character(score_id_new)) 

```


# Inter Rater Reliability: Number of Scores per Article

```{r irr}

# compute number of scores per article
# for data_r1
numbers_r1 <- data_r1 %>%
  group_by(article_id) %>%
  summarize(score_count_r1 = max(score_count)) 
# for data_r2
numbers_r2 <- data_r2 %>% 
  group_by(article_id) %>%
  summarize(score_count_r2 = max(score_count))

# join tables of data_r1 and data_r2
numbers <- full_join(numbers_r1, numbers_r2,
                     by = "article_id") 

# Krippendorff's alpha for number of scores per article
# create wide format
numbers_wide <- numbers %>%
  gather(key = "Variable", value = "Value", -article_id) %>%
  spread(key = "article_id", value = "Value") %>%
  select(-Variable)

# compute Krippendorff's alpha 
kripp_numbers <- numbers_wide %>%
  as.matrix
kripp.alpha(kripp_numbers, "ratio")

# returns krippendorffs alpha for one bootstrap resample
alpha_fun <- function(data, indices) {
  d <- data[indices, ]
  d_transposed <- t(d)
  return(kripp.alpha(d_transposed, "ratio")$value)
}

# seed 
set.seed(42)

# bootstrapping with 1000 iterations
boot_alpha_numbers <- boot(
  data = t(kripp_numbers),
  statistic = alpha_fun,
  R = 1000
)

# calculate SE and CI
alpha_se_numbers <- sd(boot_alpha_numbers$t)  
alpha_ci_numbers <- boot.ci(boot_alpha_numbers, type = "perc")$percent[4:5]  

# print se and ci
print(alpha_se_numbers)
print(alpha_ci_numbers)

```

# Inter Rater Reliability: Distinct Scores per Article

```{r irr3}

# a function to pair the scores 
pair_scores <- function(scores1, scores2) {
  common <- intersect(scores1, scores2)
  s1_only <- setdiff(scores1, common)
  s2_only <- setdiff(scores2, common)
  # pair common scores 
  paired_common <- tibble(r1 = common, r2 = common)
  # fill missings with NA 
  len_diff <- abs(length(s1_only) - length(s2_only))
  if (length(s1_only) < length(s2_only)) {
    s1_only <- c(s1_only, rep(NA, len_diff))
  } else {
    s2_only <- c(s2_only, rep(NA, len_diff))
  }
  paired_remaining <- tibble(r1 = s1_only, r2 = s2_only)
  bind_rows(paired_common, paired_remaining)
}

# create data frames with selected variables
df_r1 <- data_r1 %>% 
  select(article_id, score_value = score_id_new) %>% 
  distinct()
df_r2 <- data_r2 %>% 
  select(article_id, score_value = score_id_new) %>% 
  distinct()

# object with unique article_id values
all_ids <- unique(data_r1$article_id)

# get a list with paired scores 
paired_scores_list <- map(all_ids, function(id) {
  s1 <- df_r1 %>% filter(article_id == id) %>% pull(score_value)
  s2 <- df_r2 %>% filter(article_id == id) %>% pull(score_value)
  pair_scores(s1, s2) %>% mutate(article_id = id)
})

# create data frame with paired scores 
# In order not to overestimate the IRR by eliminating NA values, these are named as a score
paired_scores <- bind_rows(paired_scores_list) %>%
  mutate(
    r1 = ifelse(is.na(r1), "NA-score", r1),
    r2 = ifelse(is.na(r2), "NA-score", r2)
  )

# create rating matrix for cohens kappa
ratings_matrix <- paired_scores %>% 
  select(r1, r2) %>% 
  as.matrix() 

# compute kappa 
kappa_results <- kappa2(ratings_matrix)

# print kappa
print(kappa_results)

# returns cohens kappa for one bootstrap resample
kappa_fun <- function(data, indices) {
  d <- data[indices, ]
  kappa2(as.matrix(d))$value
}

# bootstrapping with 1000 iterations
boot_kappa <- boot(
  data = ratings_matrix,
  statistic = kappa_fun,
  R = 1000
)

# calculate SE and CI
kappa_se <- sd(boot_kappa$t)
kappa_ci <- boot.ci(boot_kappa, type = "perc")$percent[4:5]

# print se and ci
print(kappa_se)
print(kappa_ci)

```

# Inter Rater Reliability: Occurrence of Each Distinct Score in the Whole Sample

```{r irr2}

# add variable for frequency of each score ID
# for data_r1
score_freq_r1 <- data_r1 %>%
  group_by(score_id_new) %>%
  mutate(frequency_r1 = n()) %>%
  distinct(score_id_new, .keep_all = TRUE) %>%
  select(score_id_new, frequency_r1)

# for data_r2
score_freq_r2 <- data_r2 %>% 
  group_by(score_id_new) %>%
  mutate(frequency_r2 = n()) %>% 
  distinct(score_id, .keep_all = TRUE) %>%
  select(score_id_new, frequency_r2)

# join tables
score_freq <- full_join(score_freq_r1, score_freq_r2,
                        by = "score_id_new") %>%
  distinct(score_id_new, .keep_all = TRUE) 

# Krippendorff's alpha for occurrence of each score in the whole sample
# create wide format
score_freq_wide <- score_freq %>%
  gather(key = "Variable", value = "Value", -score_id_new) %>%
  spread(key = "score_id_new", value = "Value") %>%
  select(-Variable)

# compute Krippendorff's alpha 
kripp_score_freq <- score_freq_wide %>%
  as.matrix
kripp.alpha(kripp_score_freq, "ratio")

# bootstrapping with 1000 iterations
boot_alpha_freq <- boot(
  data = t(kripp_score_freq),
  statistic = alpha_fun,
  R = 1000
)

# calculate SE and CI
alpha_se_freq <- sd(boot_alpha_freq$t, na.rm = TRUE) 
alpha_ci_freq <- boot.ci(boot_alpha_freq, type = "perc")$percent[4:5] 

# print se and ci
print(alpha_se_freq)
print(alpha_ci_freq)

```

# Text in Manuscript

The alpha and kappa values above are those reported in the manuscript under the "Interrater reliability" heading.

# Session Info

```{r session-info}
sessionInfo()
```
