---
title: "Inter Rater Reliability for Procedural Variations"
author: "Jan-Paul Ries"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_download: true
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup}
#| include: false

# load knitr 
library(knitr)

# set knitr chunk options 
opts_chunk$set(echo = TRUE,
                message = FALSE,
                warning = FALSE)
```

## Load Packages

```{r}

library(readxl)
library(dplyr)
library(irr)
library(knitr)
library(kableExtra)
library(janitor)

```

# Read data 

```{r}

df_r1 <- readxl::read_xlsx("../../data/raw/interrater reliability/codebook - rater 1/r1_procedure_aligned.xlsx")
df_r2 <- readxl::read_xlsx("../../data/raw/interrater reliability/codebook - rater 2/r2_procedure_aligned.xlsx")

```

# Wrangle

Combine files with correct column order

```{r}

df_combined <- bind_cols(
  df_r1$article_id, df_r2$article_id,
  df_r1$authors, df_r2$authors,
  df_r1$year, df_r2$year,
  df_r1$order_number, df_r2$order_number,
  df_r1$sample_size, df_r2$sample_size,
  df_r1$variant, df_r2$variant,
  df_r1$protocol, df_r2$protocol,
  df_r1$goal, df_r2$goal,
  df_r1$instructions, df_r2$instructions,
  df_r1$feedback, df_r2$feedback,
  df_r1$incentives, df_r2$incentives,
  df_r1$currency, df_r2$currency,
  df_r1$starting_capital, df_r2$starting_capital,
  df_r1$reward_adv, df_r2$reward_adv,
  df_r1$reward_disadv, df_r2$reward_disadv,
  df_r1$gain1, df_r2$gain1,
  df_r1$gain2, df_r2$gain2,
  df_r1$loss1, df_r2$loss1,
  df_r1$loss2, df_r2$loss2,
  df_r1$gain_int1, df_r2$gain_int1,
  df_r1$gain_int2, df_r2$gain_int2,
  df_r1$loss_int1, df_r2$loss_int1,
  df_r1$loss_int2, df_r2$loss_int2,
  df_r1$freq_adv1, df_r2$freq_adv1,
  df_r1$freq_adv2, df_r2$freq_adv2,
  df_r1$freq_disadv1, df_r2$freq_disadv1,
  df_r1$freq_disadv2, df_r2$freq_disadv2,
  df_r1$n_blocks, df_r2$n_blocks,
  df_r1$n_trials, df_r2$n_trials,
  df_r1$notes_procedure, df_r2$notes_procedure,
  df_r1$rater_ID, df_r2$rater_ID
)

```

# Calculate Cohen's Kappa

```{r}

# Create an empty data frame to store Kappa values
kappa_results <- data.frame(Column1 = numeric(),
                               Column2 = numeric(),
                               Kappa = numeric(),
                               stringsAsFactors = FALSE)

# Loop through column pairs and calculate Cohen's Kappa
for (i in seq(1, 61, by = 2)) {
  kappa_result <- kappa2(df_combined[, c(i, i+1)])
  
  # Append results to the data frame
  kappa_results <- rbind(kappa_results, 
                            c(Column1 = i, Column2 = i+1, Kappa = kappa_result$value))
}

# add column containing column names
colnames <- colnames(df_r1)
kappa_results$col_name <- colnames

# rename column holding Kappa values
kappa_results <- kappa_results %>%
  rename(kappa = X1.1)

# delete unnecessary columns 
kappa_results <- kappa_results %>%
  dplyr::select(col_name, kappa)  

# delete unnecessary rows 
kappa_results <- kappa_results %>%
  filter(col_name != "article_id",
         col_name != "authors",
         col_name != "year",
         col_name != "order_number",
         col_name != "notes_procedure",
         col_name != "rater_ID")


```

# Tables

```{r}

kappa_results |>
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

kappa_results |>
  summarize(mean_kappa = mean(kappa),
            sd_kappa = sd(kappa),
            min_kappa = min(kappa),
            max_kappa = max(kappa)) |>
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

# Text in manuscript

The above results produce the results reported in the manuscript for the "Interrater reliabilty" heading. i.e., 

"We computed the interrater reliability for all variables in the procedure codebook, yielding a mean Cohen’s κ = .70 (SD = .10), with a range from .44 to .90."

# How do the kappa values correspond to the indeterminability rates?
```{r}

# load data 
data <- read_xlsx("../../data/raw/metamethod review/master coding/codebook_procedure_master.xlsx") 

# create data frame with relevant variables
data_ind <- data %>%
  select(sample_size, variant, protocol,
         goal, instructions, feedback, 
         incentives, currency, starting_capital, 
         contains("reward"), contains("1"), contains("2"), 
         n_blocks, n_trials)

# function: count and calculate percentages for certain values of a subset of variables
value_counts <- function(data, value_to_count, columns = NULL) {
  if (is.null(columns)) {
    columns <- colnames(data)
  } else {
    columns <- intersect(columns, colnames(data))
  }
  results <- list()
  for (col in columns) {
    count <- sum(data[[col]] == value_to_count, na.rm = TRUE)
    percentage <- round((count / nrow(data)) * 100)
    results[[col]] <- c("count" = count, "percentage" = percentage)
  }
  summary_table <- do.call(rbind, results) %>%
    as.data.frame(stringsAsFactors = FALSE) %>%
    tibble::rownames_to_column("variable")
  return(summary_table)
}

# count how often variable is set to indeterminable
indeterminables <- value_counts(data_ind, "indeterminable") 

# join data frame with kappa values
joined <- kappa_results %>%
  rename(variable = col_name) %>%
  left_join(indeterminables, by = "variable") %>%
  select(variable, count, percentage, kappa) %>%
  arrange(desc(kappa)) %>%
  mutate_if(is.numeric, round_half_up, digits = 2)

# print data frame with frequency of indeterminable values 
joined |>
  kable() |>
  kable_classic(full_width = FALSE)

```

