---
title: "Inter Rater Reliability for Procedural Variations"
author: "Jan-Paul Ries"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_download: true
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup}
#| include: false

# load knitr 
library(knitr)

# set knitr chunk options 
opts_chunk$set(echo = TRUE,
                message = FALSE,
                warning = FALSE)
```

## Load Packages

```{r}

library(readxl)
library(dplyr)
library(irr)
library(knitr)
library(kableExtra)
library(janitor)

```

# Read data 

```{r}

df_r1 <- readxl::read_xlsx("../../data/raw/interrater reliability/codebook - rater 1/r1_procedure_aligned.xlsx")
df_r2 <- readxl::read_xlsx("../../data/raw/interrater reliability/codebook - rater 2/r2_procedure_aligned.xlsx")

```

# Wrangle

Combine files with correct column order

```{r}

df_combined <- bind_cols(
  df_r1$article_id, df_r2$article_id,
  df_r1$authors, df_r2$authors,
  df_r1$year, df_r2$year,
  df_r1$order_number, df_r2$order_number,
  df_r1$sample_size, df_r2$sample_size,
  df_r1$variant, df_r2$variant,
  df_r1$protocol, df_r2$protocol,
  df_r1$goal, df_r2$goal,
  df_r1$instructions, df_r2$instructions,
  df_r1$feedback, df_r2$feedback,
  df_r1$incentives, df_r2$incentives,
  df_r1$currency, df_r2$currency,
  df_r1$starting_capital, df_r2$starting_capital,
  df_r1$reward_adv, df_r2$reward_adv,
  df_r1$reward_disadv, df_r2$reward_disadv,
  df_r1$gain1, df_r2$gain1,
  df_r1$gain2, df_r2$gain2,
  df_r1$loss1, df_r2$loss1,
  df_r1$loss2, df_r2$loss2,
  df_r1$gain_int1, df_r2$gain_int1,
  df_r1$gain_int2, df_r2$gain_int2,
  df_r1$loss_int1, df_r2$loss_int1,
  df_r1$loss_int2, df_r2$loss_int2,
  df_r1$freq_adv1, df_r2$freq_adv1,
  df_r1$freq_adv2, df_r2$freq_adv2,
  df_r1$freq_disadv1, df_r2$freq_disadv1,
  df_r1$freq_disadv2, df_r2$freq_disadv2,
  df_r1$n_blocks, df_r2$n_blocks,
  df_r1$n_trials, df_r2$n_trials,
  df_r1$notes_procedure, df_r2$notes_procedure,
  df_r1$rater_ID, df_r2$rater_ID
)

```

# Calculate Cohen's Kappa

```{r}

# Create an empty data frame to store Kappa values
kappa_results <- data.frame(Column1 = numeric(),
                               Column2 = numeric(),
                               Kappa = numeric(),
                               stringsAsFactors = FALSE)

# Loop through column pairs and calculate Cohen's Kappa
for (i in seq(1, 61, by = 2)) {
  kappa_result <- kappa2(df_combined[, c(i, i+1)])
  
  # Append results to the data frame
  kappa_results <- rbind(kappa_results, 
                            c(Column1 = i, Column2 = i+1, Kappa = kappa_result$value))
}

# add column containing column names
colnames <- colnames(df_r1)
kappa_results$col_name <- colnames

# rename column holding Kappa values
kappa_results <- kappa_results %>%
  rename(kappa = X1.1)

# delete unnecessary columns 
kappa_results <- kappa_results %>%
  dplyr::select(col_name, kappa)  

# delete unnecessary rows 
kappa_results <- kappa_results %>%
  filter(col_name != "article_id",
         col_name != "authors",
         col_name != "year",
         col_name != "order_number",
         col_name != "notes_procedure",
         col_name != "rater_ID")


```

# Tables

```{r}

kappa_results |>
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

kappa_results |>
  summarize(mean_kappa = mean(kappa),
            sd_kappa = sd(kappa),
            min_kappa = min(kappa),
            max_kappa = max(kappa)) |>
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  kable_classic(full_width = FALSE)

```

# Text in manuscript

The above results produce the results reported in the manuscript for the "Interrater reliabilty" heading. i.e., 

"We computed the interrater reliability for all variables in the procedure codebook, yielding a mean Cohen’s κ = .70 (SD = .10), with a range from .44 to .90."


